---
title: "Introduction to gigs"
description: >
  Learn how to get started with the basics of gigs.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gigs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`gigs` is designed to make working with outputs from either the INTERGROWTH-21<sup>st</sup> project and World Health
Organisation Child Growth Standards as easy as possible. It also provides functions for easy and reproducible
classification of newborn and infant growth.

Which functions you want to use depend on whether you're **classifying observations**, or **converting between values
and z-scores/percentiles**. Let's start by loading the package.

```{r setup}
library(gigs)
```

## Converting between z-scores/percentiles and values

### From values to z-scores/percentiles
Much of the time, you'll want to know where a measurement for a given case lies relative to 'expected' values for that
case. This is where the `value2zscore` and `value2percentile` functions come in:

#### Example 1a: Converting newborn weights
```{r example_1a}
df_newborn_weights <- data.frame(
        weight_kg = c(0.91, 3.33, 3.03, 3.35, 3.47, 1.90),
        gest_age = c(183, 280, 293, 259, 277, 231),
        sex = c(rep("M", 3), rep("F", 3))
)
df_newborn_weights$weight_p <- round(ig_nbs_value2percentile(y = df_newborn_weights$weight_kg,
                                                             gest_age = df_newborn_weights$gest_age,
                                                             sex = df_newborn_weights$sex,
                                                             acronym = "wfga"),
                                     digits = 2)

df_newborn_weights$weight_z <- round(ig_nbs_value2zscore(y = df_newborn_weights$weight_kg,
                                                         gest_age = df_newborn_weights$gest_age,
                                                         sex = df_newborn_weights$sex,
                                                         acronym = "wfga"),
                                     digits = 2)
head(df_newborn_weights)
```

Given that the outputs of these functions are vectors, gain functions integrate easily with `dplyr` verbs.

### From z-scores/percentiles to values

Alternatively, you may wish to calculate reference values for median head circumference in female newborns for
gestational ages across a range of gestational ages. You can do this with the relevant `percentile2value` function:
```{r conversion_hcfga}
ig_nbs_hcfga_percentile2value(p = 0.5,
                              gest_age = 266:272,
                              sex = "M") |>
        round(digits = 1)
```

## Classifying z-scores/percentiles and values

Sometimes it is more valuable to classify observed z-scores/percentiles than to work with the raw values, for example
when evaluating size for gestational age in newborns. The classification functions offered by this package allow the
user to easily classify growth status:

### Classifying size for gestational age
```{r example_2}
df_newborn_weights$classification <- classify_sga(weight_kg = df_newborn_weights$weight_kg,
                                                  gest_age = df_newborn_weights$gest_age,
                                                  sex = df_newborn_weights$sex)
head(df_newborn_weights)
```
