---
title: "Introduction to gigs"
description: >
  Learn how to get started with the basics of gigs.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gigs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

gigs is designed to make working with outputs from either the
INTERGROWTH-21<sup>st</sup> project and World Health Organisation Child Growth
Standards as easy as possible. It also provides functions for easy and
reproducible classification of fetal, newborn, and infant growth.

Which functions you want to use depend on whether you're **converting between
values and z-scores/centiles**, or **classifying observations**. Let's start
by loading the package.

```{r setup}
library(gigs)
```

## Converting between z-scores/centiles and values

### From values to z-scores/centiles
Much of the time, you'll want to know where a measurement for a given
observation lies relative to 'expected' values for that observation. This is
where the `value2zscore` and `value2centile` functions are useful:

#### Example 1a: Converting newborn weights
```{r example_1a}
df_newborn_weights <- data.frame(
        weight_kg = c(0.91, 3.33, 2.75, 3.35, 3.47, 1.90),
        gest_days = c(183, 280, 293, 259, 277, 231),
        sex = c(rep("M", 3), rep("F", 3))
)
# Use the `acronym` parameter to specify the standard in use...
df_newborn_weights$weight_p <- with(
  df_newborn_weights,
  ig_nbs_value2centile(y = weight_kg,
                       gest_days = gest_days,
                       sex = sex,
                       acronym = "wfga")) |>
  round(digits = 2)

# Or put the acronym in the function call to use a standard-specific function
df_newborn_weights$weight_z <- with(
  df_newborn_weights,
  ig_nbs_wfga_value2zscore(weight_kg = weight_kg,
                           gest_days = gest_days,
                           sex = sex)) |>
  round(digits = 2)
df_newborn_weights
```

Given that the outputs of these functions are vectors, gigs functions integrate
easily with [dplyr](https://dplyr.tidyverse.org/) verbs.

### From z-scores/centiles to values

For data simulation or for reconstructing centile curves, it can be useful to
calculate an expected measurement for some z-score or centile. This is where the
`zscore2value` and `centile2value` functions come in. For example, you could
generate calculate median head circumferences for female newborns across a range
of gestational ages:
```{r conversion_hcfga}
ig_nbs_hcfga_centile2value(p = 0.5,
                           gest_days = 266:272,
                           sex = "F") |>
        round(digits = 1)
```

## Classifying observations
In addition to generating continuous z-scores and centiles, you may wish to
classify attained growth. The classification functions offered by this package
allow the user to easily classify growth status:

### Size for gestational age
The `classify_sfga()` function offers size for GA classifications based on
cut-offs from the research literature. Typically, the 10<sup>th</sup> and
90<sup>th</sup> centiles are used as the size for GA cut-offs, but some
publications also specify severely small-for-GA newborns as those below the
third weight-for-GA centile.
```{r example_sga}
df_newborn_weights <- classify_sfga(.data = df_newborn_weights,
                                    weight_kg = weight_kg,
                                    gest_days = gest_days,
                                    sex = sex)
df_newborn_weights
```

### Small vulnerable newborns
The `classify_svn()` function offers small, vulnerable newborn classifications,
based on the framework set out in the Lancet's 2023 [Small Vulnerable Newborn
Series](https://www.thelancet.com/infographics-do/small-vulnerable-newborns-2023).
Essentially, `classify_svn()` stratifies results from `classify_sfga()` by
whether newborns are term or preterm:
```{r example_svn}
df_newborn_weights <- classify_svn(.data = df_newborn_weights[, c(1:4, 6)],
                                   weight_kg = weight_kg,
                                   gest_days = gest_days,
                                   sex = sex)
df_newborn_weights
```

### Stunting
The `classify_stunting()` function offers stunting (length/height-for-age)
classifications based on cut-offs from the research literature. The output is a
factor with levels corresponding to different degrees of stunting:
```{r example_stunting}
stunting_df <- data.frame(lenht_cm = c(52.2, 75.4, 63.1),
                          age_days = c(357, 375, 250),
                          gest_days = c(196, 287, 266),
                          sex = c("M", "M", "F"))
stunting_df <- classify_stunting(stunting_df,
                                 lenht_cm,
                                 age_days,
                                 gest_days,
                                 sex)
stunting_df
```

### Wasting
The `classify_wasting()` function offers wasting (weight-for-length)
classifications based on cut-offs from the research literature. The output is a
factor with levels corresponding to different degrees of wasting:
```{r example_wasting}
wasting_df <- data.frame(weight_kg = c(5.75, 2.18, 3.00, 6.75),
                         lenht_cm = c(64.6, 46.6, 50.0, 80.1),
                         gest_days = c(268, 247, 250, 278),
                         age_days = c(45, 33, 230, 278),
                         sex =  c("F", "M", "F", "M"))

wasting_df <- classify_wasting(.data = wasting_df,
                               weight_kg,
                               lenht_cm,
                               age_days,
                               gest_days,
                               sex)

wasting_df
```

### Weight-for-age
The `classify_wfa()` function offers weight-for-age classifications based on
cut-offs from the research literature. The output is a factor with levels
corresponding to different degrees of underweight and overweight:
```{r example_weight_for_age}
wfa_df <- data.frame(weight_kg = c(7.2, 4.5, 9.1, 24),
                     age_days = c(401, 185, 101, 607),
                     gest_days = 7 * c(27, 36, 40, 41),
                     sex = c("F", "M", "F", "M"))

wfa_df <- classify_wfa(wfa_df, weight_kg, age_days, gest_days, sex)
wfa_df
```

### Head size
The `classify_headsize()` function offers classification of head size based on
cut-offs from the research literature. The output is a factor with levels
corresponding to different degrees of microcephaly and macrocephaly:

```{r example_headsize}
headsize_df <- data.frame(headcirc_cm = c(48, 42, 40, 43),
                          age_days = c(401, 185, 101, 607),
                          gest_days = 7 * c(27, 36, 40, 41),
                          sex = c("F", "M", "F", "M"))

headsize_df <- classify_headsize(.data = headsize_df,
                                 headcirc_cm,
                                 age_days,
                                 gest_days,
                                 sex)
headsize_df
```