---
title: "Introduction to gigs"
description: >
  Learn how to get started with the basics of gigs.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gigs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

gigs is designed to make working with outputs from either the
INTERGROWTH-21<sup>st</sup> project and World Health Organisation Child Growth
Standards as easy as possible. It also provides functions for easy and
reproducible classification of newborn and infant growth.

Which functions you want to use depend on whether you're **converting between
values and z-scores/centiles**, or **classifying observations**. Let's start
by loading the package.

```{r setup}
library(gigs)
```

## Converting between z-scores/centiles and values

### From values to z-scores/centiles
Much of the time, you'll want to know where a measurement for a given
observation lies relative to 'expected' values for that observation. This is
where the `value2zscore` and `value2centile` functions are useful:

#### Example 1a: Converting newborn weights
```{r example_1a}
df_newborn_weights <- data.frame(
        weight_kg = c(0.91, 3.33, 2.75, 3.35, 3.47, 1.90),
        gest_days = c(183, 280, 293, 259, 277, 231),
        sex = c(rep("M", 3), rep("F", 3))
)
df_newborn_weights$weight_p <-
  with(df_newborn_weights,
       ig_nbs_value2centile(y = weight_kg,
                               gest_days = gest_days,
                               sex = sex,
                               acronym = "wfga")) |>
        round(digits = 2)

df_newborn_weights$weight_z <-
    with(df_newborn_weights,
       ig_nbs_value2zscore(y = weight_kg,
                           gest_days = gest_days,
                           sex = sex,
                           acronym = "wfga")) |>
        round(digits = 2)
df_newborn_weights
```

Given that the outputs of these functions are vectors, gigs functions integrate
easily with [dplyr](https://dplyr.tidyverse.org/) verbs. We only avoid using
them here to reduce the package dependencies for gigs.

### From z-scores/centiles to values
Alternatively, you may wish to calculate reference values for median head
circumference in female newborns across a range of gestational ages. You can do
this with the relevant `centile2value` function:
```{r conversion_hcfga}
ig_nbs_hcfga_centile2value(p = 0.5,
                           gest_days = 266:272,
                           sex = "M") |>
        round(digits = 1)
```

## Classifying observations
Sometimes it is more valuable to classify observed z-scores/centiles than to
work with the raw values, for example when evaluating size for gestational age
in newborns. The classification functions offered by this package allow the user
to easily classify growth status:

### Size for gestational age
The `classify_sga()` function offers size for GA classifications based on
cut-offs from the research literature. Typically, the 10<sup>th</sup> and
90<sup>th</sup> centiles are used as the size for GA cut-offs, but some
publications also specify severely small-for-GA newborns as those below the
third weight-for-GA centile. Whether you do this depends on whether you set the
`severe` parameter in `classify_sga()` to `TRUE`:
```{r example_sga}
df_newborn_weights$sga <- with(
        df_newborn_weights,
        classify_sfga(weight_kg = weight_kg, gest_days = gest_days, sex = sex))
df_newborn_weights$sga_severe <- with(
        df_newborn_weights,
        classify_sfga(weight_kg = weight_kg,
                      gest_days = gest_days,
                      sex = sex,
                      severe = TRUE))
df_newborn_weights
```

### Small vulnerable newborns
The `classify_svn()` function offers small, vulnerable newborn classifications,
based on the framework set out in the Lancet's 2023 Small Vulnerable Newborn
Series. Essentially, this stratifies results from `classify_sga()` by whether
newborns are term or preterm:
```{r example_svn}
df_newborn_weights$svn <- with(
        df_newborn_weights,
        classify_svn(weight_kg = weight_kg, gest_days = gest_days, sex = sex))
df_newborn_weights
```

### Stunting
The `classify_stunting()` function offers stunting (length/height-for-age)
classifications based on cut-offs from the research literature. The output is a
factor with levels corresponding to different degrees of stunting:
```{r example_stunting}
classify_stunting(
  lenht_cm = c(52.2, 75.4, 63.1),
  age_days = c(357, 375, 250),
  gest_days = c(196, 287, 266),
  sex = c("M", "M", "F")
)
```

You can also specify for `classify_stunting()` to classify likely
outlier/implausible values by setting the `outlier` parameter to `TRUE`:
```{r example_stunting_outliers}
classify_stunting(
  lenht_cm = c(52.2, 75.4, 63.1),
  age_days = c(357, 375, 250),
  gest_days = c(196, 287, 266),
  sex = c("M", "M", "F"),
  outliers = TRUE
)
```

### Wasting
The `classify_wasting()` function offers wasting (weight-for-length)
classifications based on cut-offs from the research literature. The output is a
factor with levels corresponding to different degrees of wasting:
```{r example_wasting}
classify_wasting(
  weight_kg = c(5.75, 2.18, 3.00, 6.75),
  lenht_cm = c(67.7, 46.6, 50.0, 80.1),
  gest_days = c(268, 247, 250, 278),
  age_days = c(45, 33, 230, 278),
  sex =  c("F", "M", "F", "M")
)
```

You can also specify for `classify_wasting()` to classify likely
outlier/implausible values by setting the `outlier` parameter to `TRUE`:
```{r example_wasting_outliers}
classify_wasting(
  weight_kg = c(5.75, 2.18, 3.00, 6.75),
  lenht_cm = c(67.7, 46.6, 50.0, 80.1),
  gest_days = c(268, 247, 250, 278),
  age_days = c(45, 33, 230, 278),
  sex =  c("F", "M", "F", "M"),
  outliers = TRUE
)
```

### Weight-for-age
The `classify_wfa()` function offers weight-for-age classifications based on
cut-offs from the research literature. The output is a factor with levels
corresponding to different degrees of underweight and overweight:
```{r example_weight_for_age}
classify_wfa(
  weight_kg = c(7.2, 4.5, 9.1, 24),
  age_days = c(401, 185, 101, 607),
  gest_days = 7 * c(27, 36, 40, 41),
  sex = c("F", "M", "F", "M")
)
```

You can also specify for `classify_wasting()` to classify likely
outlier/implausible values by setting the `outlier` parameter to `TRUE`:
```{r example_weight_for_age_outliers}
classify_wfa(
  weight_kg = c(7.2, 4.5, 9.1, 24),
  age_days = c(401, 185, 101, 607),
  gest_days = 7 * c(27, 36, 40, 41),
  sex = c("F", "M", "F", "M"),
  outliers = TRUE
)
```
