---
title: "Benchmarking gigs against other software packages"
description: >
  A comparison of gigs' performance against other growth analysis packages.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dpi = 120, message = FALSE, warning = FALSE,
  dev.args = list(png = list(type = "cairo"))
)

# Set ggplot theming for consistent plots
ggplot2::theme_set(ggplot2::theme_bw())
intergrowth21st <- knitr::asis_output(x = "INTERGROWTH-21<sup>st</sup>")
```

```{r srr-tags, eval = FALSE, echo = FALSE}
#' @srrstats {1.5, 1.6} This article contains code for both performance and
#'   accuracy claims when comparing this package with other software packages.
```

```{r load_timings_rda, echo = FALSE}
load(file = file.path("benchmarking.rda"))
```

# Rationale
This short article describes the performance of gigs relative to a
non-exhaustive group of R and non-R packages designed to implement child growth
standards.

| Package name                                                                 | Language | On CRAN? |
|------------------------------------------------------------------------------|:---------|:---------|
| [anthro](https://cran.r-project.org/web/packages/anthro/index.html)          | R        | Yes      |
| [childsds](https://cran.r-project.org/web/packages/childsds/index.html)      | R        | Yes      |
| [gigs](https://www.github.com/lshtm-gigs/gigs/)                              | R        | No       |
| [ki-tools/growthstandards](https://www.github.com/ki-tools/growthstandards/) | R        | No       |
| [gigs](https://www.github.com/lshtm-gigs/gigs-stata/)                        | Stata    | No       |
| [zanthro](https://journals.sagepub.com/doi/epdf/10.1177/1536867X1301300211)  | Stata    | No       |

Thus far, there is no comprehensive benchmark comparing these different
packages. This short article will compare the speed of each package from 1 to
100000 inputs, checking how fast each package can convert weight values to
z-scores in different growth standards.

We performed these benchmarks on R version
`r session_info[[1]][["major"]]`.`r session_info[[1]][["minor"]]`, using a
Windows 10 system with a Ryzen 7 3700X processor and 16GB of DDR4 RAM. The Stata
benchmarks were run in Stata 18.0 (revision 15 May 2023) on the same system,
using the [benchmark](https://github.com/mcaceresb/stata-benchmark) package for Stata.

Other R packages can be used to analyse growth data with international
standards, but have limitations which are not present in gigs. There are also
software packages external to R which implement these standards. The table
below describes these packages, and to what extent they have implemented
functions that let users convert anthropometric measurements to
z-scores/centiles in each set of standards implemented in gigs - the WHO Child
Growth standards, `r intergrowth21st` Newborn Size standards (including Very
Preterm), and the `r intergrowth21st` Postnatal Growth standards for preterm
infants. A green tick (✅) indicates that a package contains functions for all
growth standards within a set, a red cross (❌) indicates that a package has no
functions for a set of growth standards, and an exclamation mark (❕) indicates
that some standards in a set are implemented, but not others.

```{r arrows, echo = FALSE}
ig21st <- knitr::asis_output(x = "IG-21<sup>st</sup>")

larr <- knitr::asis_output(x = "&larr;")
rarr <- knitr::asis_output(x = "&rarr;")
harr <- knitr::asis_output(x = "&harr;")
yes <- knitr::asis_output(x = "✅")
no <- knitr::asis_output(x = "❌")
half <- knitr::asis_output(x = "❕")

anthro <- knitr::asis_output(x = "[anthro](https://cran.r-project.org/web/packages/anthro/index.html)")
childsds <- knitr::asis_output(x = "[childsds](https://cran.r-project.org/web/packages/childsds/index.html)")
gigs_r <- knitr::asis_output(x = "[gigs](https://www.github.com/lshtm-gigs/gigs/)")
growthstans <- knitr::asis_output(x = "[ki-tools/growthstandards](https://www.github.com/ki-tools/growthstandards/)")
gigs_stata <- knitr::asis_output(x = "[gigs](https://www.github.com/lshtm-gigs/gigs-stata/) (Stata)")
zanthro <- knitr::asis_output(x = "[zanthro](https://journals.sagepub.com/doi/epdf/10.1177/1536867X1301300211) (Stata)")
intergrowth <- knitr::asis_output(x = "[nutriverse/intergrowth](https://github.com/nutriverse/intergrowth/)")
```

| Software        | Platform | WHO (0-5 years) | `r ig21st` NBS | `r ig21st` PNG | `r ig21st` Fetal | Functionality                     |
|-----------------|----------|-----------------|----------------|----------------|------------------|-----------------------------------|
| `r gigs_r`      | R        | `r yes`         | `r yes`        | `r yes`        | `r yes`          | Values `r harr` z-scores/centiles |
| `r anthro`      | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores          |
| `r childsds`    | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores/centiles |
| `r growthstans` | R        | `r yes`         | `r yes`        | `r half`       | `r half`         | Values `r harr` z-scores/centiles |
| `r intergrowth` | R        | `r no`          | `r no`         | `r no`         | `r half`         | Values `r rarr` z-scores/centiles |
| `r gigs_stata`  | Stata    | `r yes`         | `r yes`        | `r yes`        | `r yes`          | Values `r harr` z-scores/centiles |
| `r zanthro`     | Stata    | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores/centiles |

# Shared benchmark code

```{r load_gigs, eval = FALSE}
library(gigs)
```

In this vignette, we perform benchmark performance in each a growth standard
from each set of growth standards implemented in gigs. These are the:

* WHO child growth standards (0-5 years)
* `r intergrowth21st` Newborn Size standards
* `r intergrowth21st` Postnatal Growth standards
* `r intergrowth21st` Fetal standards

The benchmarks utilise pre-generated datasets for growth standard. The z-scores,
`x` variable and sexes for each row are generated (pseudo)randomly. Because the
gigs test suite validates its functions against published growth curve data from
the WHO and `r intergrowth21st` Project, we use gigs to generate expected
measurements for each observation. This dataset can then be used to compare
speeds for each package. These datasets are generated with the following
function:
```{r benchmark_dataset, echo = TRUE}
benchmark_dataset <- function(n, standard, acronym) {
  data <- rlang::eval_bare(rlang::parse_expr(paste0("gigs::", standard)))
  if (standard == "ig_fet") {
    xvars <- data[[acronym]][[1]][[1]]
  } else {
    xvars <- data[[acronym]][[1]][[1]][[1]]
  }

  withr::with_seed(seed = 154237890, code = {
      # Random z-scores around 0
      z <- rnorm(n = n)
    })
  withr::with_seed(seed = 154237890, code = {
      # X variables are non-discrete but within bounds of the desired acronym
      x <-  sample(xvars, size = n, replace = TRUE)
      x_jitter <- c(runif(n = 5, min = -1, max = 1), 0)
      x <- x + sample(x_jitter, size = n, replace = TRUE)
      x <- pmax(pmin(x, max(xvars)), min(xvars))
      # Sexes randomly sampled from male and female
      sex <- sample(c("M", "F"), size = n, replace = TRUE)
    }
  )
  gigs_fn <- rlang::eval_bare(
    rlang::parse_expr(paste0("gigs::", paste0(standard, "_zscore2value")))
  )
  # gigs_fn <- get(, rlang::ns_env("gigs"))
  out <- if (standard != "ig_fet") {
    if (standard == "ig_nbs") {
      list(z = z, gest_days = x, sex = sex, acronym = acronym)
    } else {
      list(z = z, x = x, sex = sex, acronym = acronym)
    }
  } else {
    list(z = z, x = x, acronym = acronym)
  }
  out$y <- do.call(what = gigs_fn, args = out)
  as.data.frame(out, stringsAsFactors = FALSE)
}
```

```{r hidden_helpers, echo = FALSE}
max_median <- function(bp, name) {
  round(max(bp[bp$desc == name, "median"]))
}

pkg_str <- function(pkgname) paste("R:", pkgname, packageVersion(pkg = pkgname))

plot_bp <- function(bp_r, bp_stata = NULL) {
  key_pkg_colour <- c("gs" = "#925E9FFF",
                      "gigs" = "#00468BFF",
                      "anthro" = "#AD002AFF",
                      "intergrowth" = "#FDAF91FF",
                      "childsds" = "#663300FF",
                      "stata_gigs" = "skyblue",
                      "stata_zanthro" = "darkolivegreen4")
  key_pkg_label <- c("gs" = pkg_str("growthstandards"),
                     "gigs" = pkg_str("gigs"),
                     "anthro" = pkg_str("anthro"),
                     "intergrowth" = pkg_str("intergrowth"),
                     "childsds" = pkg_str("childsds"),
                     "stata_gigs" = "Stata: gigs 0.4.0",
                     "stata_zanthro" = "Stata: zanthro 1.0.2")
  bp <- bp_r  |>
    dplyr::select(desc, input_len, median)
  if (!is.null(bp_stata)) {
    suppressMessages(suppressWarnings(bp <- dplyr::full_join(bp, bp_stata)))
  }
  bp |>
    ggplot2::ggplot(ggplot2::aes(x = input_len, y = median, colour = desc)) +
    ggplot2::geom_line() +
    ggplot2::geom_point(colour = "black") +
    ggplot2::labs(x = "Input lengths", y = "Median time taken (ms)",
                  colour = "Package") +
    ggplot2::scale_colour_manual(values = key_pkg_colour,
                                 labels = key_pkg_label) +
    ggplot2::theme_bw()
}
```

We're using the `{bench}` package throughout this vignette, specifically the
`bench::press()` function. This function lets us easily compare scaling for the
different functions available to us, as we can iterate through increasing
amounts of observations.

# WHO Growth Standards
The World Health Organisation (WHO) Child Growth Standards are included in gigs
with the various `who_gs_*`conversion functions. These standards are implemented
in a variety of other packages.

## Set up benchmark dataset

```{r ds_who_gs}
# Generate 100,000-row dataset
ac_who_gs <- "wfa"
ds_who_gs <- benchmark_dataset(n = 100000,
                               standard = "who_gs",
                               acronym = ac_who_gs)
```

The first 10 rows of this dataset look like this:
```{r show_bench_dataset_who_gs, echo = FALSE}
ds_who_gs[1:10, ]
```


## Time packages

```{r bp_who_gs, echo = TRUE, eval = FALSE}
bp_who_gs <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    selector <- seq_len(input_len)
    weight_kg <- ds_who_gs$y[selector]
    age_days <- ds_who_gs$x[selector]
    acronym <- ds_who_gs$acronym[selector]
    sex <- ds_who_gs$sex[selector]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    bm_who_gs <- bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::who_gs_value2zscore(weight_kg, age_days, sex, ac_who_gs),
      gs = growthstandards::who_value2zscore(y = weight_kg, x = age_days,
                                             y_var = "wtkg", x_var = "agedays",
                                             sex = sex_gs),
      childsds = childsds::sds(value = weight_kg,
                               age = age_days / 365.25,
                               sex = sex,
                               male = "M",
                               female = "F",
                               item = "weight",
                               ref = childsds::who.ref),
      anthro = anthro::anthro_zscores(weight = weight_kg,
                                      age = age_days,
                                      sex = sex)$zwei
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_who_gs,
  path = file.path("exclude", "statabench", "bench_ds_who_gs.dta")
)
```

In Stata, the commands are run inside a do-file which utilises the
[benchmark](https://www.github.com/mcaceresb/stata-benchmark) package for Stata. This code essentially does the same as
`bench::press()`, but for the packages we're testing in Stata.
```{stata stata_bench_code_who_gs, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_who_gs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = who_gs(y, "wfa", "v2z"), ///
		    xvar(x) sex(sex) sexcode(m=M, f=F)
}

foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_who_gs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen z_anthro = zanthro(y, wa, WHO), xvar(x) gender(sex) ///
            gencode(male=M, female=F) ageunit(day)
}
```

```{r bp_who_gs_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.009, 0.008, 0.009, 0.010, 0.012, 0.028, 0.047, 0.104, 0.198,
                0.296, 0.405)
zanthro_1_0_2 <- c(0.007, 0.008, 0.009, 0.017, 0.027, 0.105, 0.203, 0.479,
                   0.951, 1.4450, 2.046)
lens <- lengths(list(gigs_0_4_0, zanthro_1_0_2))
n_inputs <- unlist(lapply(X = lens, FUN = \(x) input_len[seq_len(x)]))
bp_who_gs_stata <-  data.frame(input_len = n_inputs,
                               median = c(gigs_0_4_0, zanthro_1_0_2) * 1000,
                               desc = c(rep("stata_gigs", lens[1]),
                                        rep("stata_zanthro", lens[2])))
rm(gigs_0_4_0, zanthro_1_0_2, lens, n_inputs, input_len)
```

## Package comparisons: timing

```{r plot_timings_who_gs, echo = FALSE}
plot_bp(bp_who_gs, bp_who_gs_stata)
```

On the whole, `anthro` is by far the slowest R package, taking around
`r max_median(bp_who_gs, "anthro") / 1000` seconds to process 100,000
inputs. This is in part because `anthro` computes results in every WHO Child
Growth standard each time `anthro::anthro_zscores()` is called, but also due to
a slower implementation of the WHO LMS procedure than the other packages.

Next slowest is the Stata package `zanthro`, which takes around
`r max_median(bp_who_gs_stata, "stata_zanthro") / 1000` seconds to compute
results in a single WHO standard. About 4 times faster than `zanthro` is `gigs`
for Stata 0.4.0, which scales more efficiently than `zanthro` and takes around
`r max_median(bp_who_gs_stata, "stata_gigs") / 1000` seconds to convert 100,000
measurements to z-scores.

Let's look at the three fastest R implementations:

```{r plot_timings_who_gs_faster, echo = FALSE}
bp_who_gs |>
  dplyr::filter(desc %in% c("gigs", "childsds", "gs")) |>
  plot_bp()
```

The `growthstandards` package was the fastest at
~`r max_median(bp_who_gs, "gs")` ms for 100,000 inputs, followed by `gigs`
(~`r max_median(bp_who_gs, "gigs")` ms) and `childsds`
(~`r max_median(bp_who_gs, "childsds")` ms).

## Package comparisons: numerical consistency
In our testing of the WHO standards, we found that the tested packages mostly
agreed with each other when provided the same inputs, except for `childsds`.

This is because the WHO Child Growth standards constrain z-scores in the outer
tails (i.e. past 3 z-scores), as their data was more sparse for these extreme
values. More information on this constraining procedure can be found in the
reports referenced in the `gigs::who_gs_value2zscore()` documentation. It
appears that `childsds` does not perform this constraining procedure, so
conversions where `abs(z_score) > 3` are computed incorrectly:
```{r discrepancies_who_gs}
discrepancies <- data.frame(z = c(-3.03, -2.97, 2.97, 3.03),
                            age_days = 0,
                            sex = "M") |>
  dplyr::mutate(
    weight_kg = gigs::who_gs_wfa_zscore2value(z, age_days, sex),
    # GIGS z-score
    z_gigs = gigs::who_gs_wfa_value2zscore(weight_kg, age_days, sex),
    # growthstandards z-score
    z_growthstandards = growthstandards::who_wtkg2zscore(
      age_days, weight_kg, "Male"
    ),
    # childsds z-score
    z_childsds = childsds::sds(
      value = weight_kg, age = age_days / 365.25,
      sex = sex, male = "M", female = "F",
      item = "weight", ref = childsds::who.ref
    )
  )
```

When we look at these z-scores, you can see that both `growthstandards` and
`gigs` correctly apply the constraining procedure; `childsds` does not. From
looking at the `anthro` source code, they also apply the constraining procedure.
```{r discrepancies_who_gs_kable, echo = FALSE}
knitr::kable(discrepancies, align = "ccccc")
```

# INTERGROWTH-21<sup>st</sup> Newborn Size standards
The `r intergrowth21st` Newborn Size standards are implemented in gigs for R and
Stata, and in the `growthstandards` package for R. Let's make a new dataset,
`ds_ig_nbs`, which we will use to benchmark these functions:

```{r ds_ig_nbs}
ac_ig_nbs <- "wfga"
ds_ig_nbs <- benchmark_dataset(n = 100000,
                               standard = "ig_nbs",
                               acronym = ac_ig_nbs)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset_ig_nbs}
ds_ig_nbs[1:10, ]
```

## Time package

```{r bp_ig_nbs, echo = TRUE, eval = FALSE}
bp_ig_nbs <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    p <- pnorm(ds_ig_nbs$z[1:input_len])
    weight_kg <- ds_ig_nbs$y[1:input_len]
    gest_days <- ds_ig_nbs$gest_days[1:input_len]
    acronym <- ds_ig_nbs$acronym[1:input_len]
    sex <- ds_ig_nbs$sex[1:input_len]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::ig_nbs_value2centile(weight_kg, gest_days, sex, ac_ig_nbs),
      gs = growthstandards::igb_value2centile(val = weight_kg,
                                              gagebrth = gest_days,
                                              sex = sex_gs, var = "wtkg")
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```
### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_nbs,
  path = file.path("exclude", "statabench", "bench_ds_ig_nbs.dta")
)
```

We can then benchmark the speed of the Stata command:
```{stata stata_bench_code_ig_nbs, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_nbs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_nbs(y, "wfga", "v2z"), ///
		    gest_days(x) sex(sex) sexcode(m=M, f=F)
}
```

```{r bp_ig_nbs_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.009, 0.004, 0.005, 0.006, 0.008, 0.027, 0.048, 0.111, 0.220,
                0.334, 0.471)
bp_ig_nbs_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_nbs, echo = FALSE}
plot_bp(bp_ig_nbs, bp_ig_nbs_stata)
```

For this set of growth standards, the Stata implementation for GIGS is the
slowest at ~`r max_median(bp_ig_nbs_stata, "stata_gigs")` ms. The
`growthstandards` and `gigs` packages are the fastest at
~`r max_median(bp_ig_nbs, "gs")` and ~`r max_median(bp_ig_nbs, "gigs")` ms,
respectively.

## Package comparisons: numerical consistency
In our testing of the `r intergrowth21st` Newborn Size standards, we found that
the implementation in `growthstandards` does not perform coefficient
interpolation for the `r intergrowth21st` Newborn Size standards which
utilise mu/sigma/nu/tau coefficients. Instead, `growthstandards` uses `round()`
to round non-integer gestational ages to the nearest value, then gets
mu/sigma/nu/tau coefficients for this rounded GA. This leads a smaller-scale
version of the z-scoring errors found by
[Kiger *et al.* (2016)](https://dx.doi.org/10.1007/s10916-015-0389-x) when not
interpolating LMS coefficients in the WHO standards.

```{r discrepancies_ig_nbs, eval = FALSE}
weight_kg <- 3
gestage <- seq(274, 278, 0.5)
gigs <- gigs::ig_nbs_value2centile(y = weight_kg, gest_days = gestage,
                                   sex = "M", acronym = "wfga")
gs <- growthstandards::igb_value2centile(val = weight_kg, gagebrth = gestage,
                                         sex = "Male", var = "wtkg")
waldo::compare(gigs, gs / 100, x_arg = "gigs", y_arg = "growthstandards")
```

# INTERGROWTH-21<sup>st</sup> Postnatal Growth standards
The `r ig21st` Postnatal Growth standards are implemented in gigs for R and
Stata, and in the `growthstandards` package. Let's make a new dataset,
`ds_ig_png`, which we will use to benchmark these functions:
```{r ds_ig_png}
ac_ig_png <- "wfa"
ds_ig_png <- benchmark_dataset(n = 100000,
                               standard = "ig_png",
                               acronym = ac_ig_png)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset4}
ds_ig_png[1:10, ]
```

## Time package

```{r bp_ig_png, echo = TRUE, eval = FALSE}
bp_ig_png <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    weight_kg <- ds_ig_png$y[1:input_len]
    pma_weeks <- ds_ig_png$x[1:input_len]
    pma_days <- (pma_weeks * 7)[1:input_len]
    acronym <- ds_ig_png$acronym[1:input_len]
    sex <- ds_ig_png$sex[1:input_len]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::ig_png_value2zscore(weight_kg, pma_weeks, sex, ac_ig_png),
      gs = growthstandards::igprepost_value2zscore(val = weight_kg,
                                                   pmagedays = pma_days,
                                                   sex = sex_gs, var = "wtkg")
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_png,
  path = file.path("exclude", "statabench", "bench_ds_ig_png.dta")
)
```

```{stata stata_bench_code_ig_png, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_png.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_png(y, "wfga", "v2z"), ///
		    xvar(x) sex(sex) sexcode(m=M, f=F)
}
```

```{r bp_ig_png_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.003, 0.002, 0.002, 0.003, 0.004, 0.010, 0.019, 0.043, 0.082,
                0.114, 0.164)
bp_ig_png_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_png, echo = FALSE}
plot_bp(bp_ig_png, bp_ig_png_stata)
```

For the `r intergrowth21st` Postnatal Growth standards, the Stata implementation
for GIGS is the slowest at ~`r max_median(bp_ig_png_stata, "stata_gigs")` ms.
The `gigs` and `growthstandards` packages are the fastest at
~`r max_median(bp_ig_png, "gs")` and ~`r max_median(bp_ig_png, "gigs")` ms,
respectively.

# INTERGROWTH-21<sup>st</sup> Fetal standards
The `r ig21st` Fetal standards are implemented in gigs for R and Stata, and in
the `growthstandards` and `intergrowth` packages (though more fully in
`intergrowth`  than in `growthstandards`, and both are missing some standards
which are included in `gigs`). Let's make a new dataset, `ds_ig_fet`, which we
will use to benchmark these functions:

```{r ds_ig_fet}
ac_ig_fet <- "ofdfga"
ds_ig_fet <- benchmark_dataset(n = 100000,
                               standard = "ig_fet",
                               acronym = ac_ig_fet)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset_ig_fet, eval = FALSE}
ds_ig_fet[1:10, ]
```

## Time package

```{r bp_ig_fet, echo = TRUE, eval = FALSE}
bp_ig_fet <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    gest_days <- ds_ig_fet$x[1:input_len]
    gest_wks <- gest_days / 7
    acronym <- ds_ig_fet$acronym[1:input_len]
    ofd_mm <- ds_ig_fet$y[1:input_len]
    ofd_cm <- ofd_mm / 10
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::ig_fet_value2zscore(ofd_mm, gest_days, ac_ig_fet),
      gs = growthstandards::igfet_value2zscore(val = ofd_cm,
                                               gagedays = gest_days,
                                               var = "ofdcm"),
      intergrowth = intergrowth::calculate_ofd_gestage(ofd = ofd_mm,
                                                       gestage = gest_wks)[[1]],
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_fet,
  path = file.path("exclude", "statabench", "bench_ds_ig_fet.dta")
)
```

```{stata stata_bench_code_ig_fet, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_fet.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_fet(y, "ofdfga", "v2z"), xvar(x)
}
```

```{r bp_ig_fet_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.002, 0.003, 0.002, 0.003, 0.004, 0.006, 0.012, 0.025, 0.046,
                0.066, 0.093)
bp_ig_fet_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_fet, echo = FALSE}
plot_bp(bp_ig_fet, bp_ig_fet_stata)
```

For the `r intergrowth21st` Fetal standard for occipito-frontal diameter, the
GIGS implementation for Stata is the slowest to analyse 100,000 inputs at
~`r max_median(bp_ig_fet_stata, "stata_gigs")` ms. GIGS for R is the fastest at
~`r max_median(bp_ig_fet, "gigs")` ms, followed by `growthstandards` at
~`r max_median(bp_ig_fet, "gs")`, and then `intergrowth` at
~`r max_median(bp_ig_fet, "intergrowth")`.

# Summary
## Timings summary

```{r save_timings, echo = FALSE, eval = FALSE}
max_median_all <- function(bp_r, bp_stata) {
  bp_both <- dplyr::bind_rows(bp_r[, c(2, 4, 15)], bp_stata)
  vapply(X = unique(bp_both$desc),
         FUN = \(desc) max_median(bp_both, desc),
         FUN.VALUE = double(1L))
}

max_len_bench <- list(
  who_gs = max_median_all(bp_who_gs, bp_who_gs_stata),
  ig_nbs = max_median_all(bp_ig_nbs, bp_ig_nbs_stata),
  ig_png = max_median_all(bp_ig_png, bp_ig_png_stata),
  ig_fet = max_median_all(bp_ig_fet, bp_ig_fet_stata)
)
```

| Software        | Platform | WHO (0-5 years) (ms)                      | `r ig21st` NBS (ms)                    | `r ig21st` PNG (ms)                    | `r ig21st` Fetal (ms)                   |
|-----------------|----------|-------------------------------------------|----------------------------------------|----------------------------------------|-----------------------------------------|
| `r gigs_r`      | R        | `r max_len_bench[[1]][["gigs"]]`          | `r max_len_bench[[2]][["gigs"]]`       | `r max_len_bench[[3]][["gigs"]]`       | `r max_len_bench[[4]][["gigs"]]`        |
| `r anthro`      | R        | `r max_len_bench[[1]][["anthro"]]`        | `r no`                                 | `r no`                                 | `r no`                                  |
| `r childsds`    | R        | `r max_len_bench[[1]][["childsds"]]`      | `r no`                                 | `r no`                                 | `r no`                                  |
| `r growthstans` | R        | `r max_len_bench[[1]][["gs"]]`            | `r  max_len_bench[[2]][["gs"]]`        | `r max_len_bench[[3]][["gs"]]`         | `r max_len_bench[[4]][["gs"]]`          |
| `r intergrowth` | R        | `r no`                                    | `r no`                                 | `r no`                                 | `r max_len_bench[[4]][["intergrowth"]]` |
| `r gigs_stata`  | Stata    | `r max_len_bench[[1]][["stata_gigs"]]`    | `r max_len_bench[[2]][["stata_gigs"]]` | `r max_len_bench[[3]][["stata_gigs"]]` | `r max_len_bench[[4]][["stata_gigs"]]`  |
| `r zanthro`     | Stata    | `r max_len_bench[[1]][["stata_zanthro"]]` | `r no`                                 | `r no`                                 | `r no`                                  |


## Why GIGS?
Overall, though the `gigs` package for R isn't always the fastest, it is only a
some milliseconds off, and offers:

  * A full complement of growth standards (i.e. no half-implemented sets of
    growth standards)
  * More accurate implementations in some cases
  * Growth categorisation functions

We think these benefits outweigh a couple of milliseconds slowdown. In Stata,
where `gigs` consistently runs faster and offers more growth standards than
`zanthro`, `gigs` is the obvious choice.

# Session information
```{r session_info1, echo = TRUE, eval = FALSE}
sessionInfo()
```

```{r session_info2, echo = FALSE, eval = FALSE}
session_info <- sessionInfo()
```

```{r session_info3, echo = FALSE, eval = TRUE}
session_info
```



[//]: # (Code to save/load all timings tables in a separate .rda file)
```{r save_benchmarking_rda, eval = FALSE, echo = FALSE}
save(file = file.path("vignettes/articles/benchmarking.rda"),
     list = ls(pattern = "^(bp)"), max_len_bench, session_info)
```