---
title: "Benchmarking gigs against other software packages"
description: >
  A comparison of gigs' performance against other growth analysis packages.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dpi = 120, message = FALSE, warning = FALSE,
  dev.args = list(png = list(type = "cairo"))
)

# Set ggplot theming for consistent plots
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r load_timings_rda, echo = FALSE}
load(file = file.path("benchmarking.rda"))
```

# Rationale
This short article describes the performance of gigs relative to a
non-exhaustive group of R and non-R packages designed to implement child growth
standards.

| Package name                                                                 | Language | On CRAN? |
|------------------------------------------------------------------------------|:---------|:---------|
| [anthro](https://cran.r-project.org/web/packages/anthro/index.html)          | R        | Yes      |
| [childsds](https://cran.r-project.org/web/packages/childsds/index.html)      | R        | Yes      |
| [gigs](https://www.github.com/lshtm-gigs/gigs/)                              | R        | No       |
| [ki-tools/growthstandards](https://www.github.com/ki-tools/growthstandards/) | R        | No       |
| [gigs](https://www.github.com/lshtm-gigs/gigs-stata/)                        | Stata    | No       |
| [zanthro](https://journals.sagepub.com/doi/epdf/10.1177/1536867X1301300211)  | Stata    | No       |

Thus far, there is no comprehensive benchmark comparing these different
packages. This short article will compare the speed of each package from 1 to
100000 inputs, checking how fast each package can convert weight values to
z-scores in the WHO Child Growth standards.

We have performed these benchmarks on a Windows 10 system running a Ryzen 7
3700X processor and 16GB of DDR4 RAM. The Stata benchmarks have been run on the
same  system in Stata 18.0, using the
[benchmark](https://github.com/mcaceresb/stata-benchmark) package for Stata.

## Set up benchmark dataset
The benchmarking and package comparisons will all use the same 100,000-row
dataset comprised of data which can be used to convert between z-scores
and centiles in the WHO weight-for-age standard. The z-scores, `x` variable
and sexes for each row are generated randomly with a pre-specified seed. As gigs
is validated against the published growth curve data from the WHO, we use it to
generate weight values in kg for each observation. This dataset can then be used
to compare speed and accuracy in conversion between values and z-scores/centiles
for the packages mentioned above.

```{r who_gs_dataset, echo = TRUE}
who_gs_dataset <- function(n, acronym = "wfa", seed = 154237890) {
  withr::with_seed(
    seed,
    code = {
      # Random z-scores around 0
      z <- rnorm(n = n)
      # X variables are non-discrete but within bounds of the desired acronym
      xvars <- gigs::who_gs_coeffs[[acronym]][[1]][, 1]
      x <-  sample(xvars, size = n, replace = TRUE)
      x_jitter <- c(runif(n = 5, min = -1, max = 1), 0)
      x <- x + sample(x_jitter, size = n, replace = TRUE)
      x <- pmax(pmin(x, max(xvars)), min(xvars))
      # Sexes randomly sampled from male and female
      sex <- sample(c("M", "F"), size = n, replace = TRUE)
    }
  )
  acronym <- rep_len(acronym, n)
  out <- data.frame(z = z, x = x, sex = sex, acronym = acronym) |>
    dplyr::mutate(
      y = gigs::who_gs_zscore2value(z = z, x = x, sex = sex, acronym = acronym)
    )
  purrr::set_names(out, c("z", "x", "sex", "acronym", "y"))
}

# Generate 100,000-row dataset
bench_dataset <- who_gs_dataset(n = 100000)
```

The first 10 rows of this dataset look like this:
```{r show_bench_dataset1, eval = FALSE}
bench_dataset[1:10, ]
```

```{r show_bench_dataset2, echo = FALSE}
knitr::kable(bench_dataset[1:10, ], align = "ccccc")
```


# Benchmark code
The `mbench_pkg()` function is used to benchmark each package over a range of
input sizes.  Each call to it produces a tabular output containing the lower
quartile, median and upper quartile timings for `pkg_expr` to operate on the
data, ranging from 1 to 100,000 inputs.

```{r iteration}
mbench_pkg <- function(pkg_expr, pkg_name) {
  n_inputs <- c(1, 10, 100, 500, 1000, 5000, 10000, 25000, 50000, 75000, 100000)
  purrr::map_dfr(.x = n_inputs,
                 .f = \(no_of_inputs) {
                   dataset <- bench_dataset[seq_len(no_of_inputs), ]
                   mbench <- microbenchmark::microbenchmark(
                     dplyr::mutate(dataset, test = eval(pkg_expr)),
                     times = 25
                   )
                   c(n_inputs = no_of_inputs,
                     median_time = summary(mbench)$median,
                     time_units = attr(summary(mbench), which = "unit"))
                 }) |>
    dplyr::mutate(n_inputs = as.integer(n_inputs),
                  median_time = as.numeric(median_time),
                  package = pkg_name)
}
```

## `anthro`
```{r bench_anthro, eval = FALSE}
anthro_timings <- mbench_pkg(
  pkg_expr = quote(anthro::anthro_zscores(sex = sex,
                                          age = x,
                                          weight = y)$zwei),
  pkg_name = paste("R: anthro", packageVersion(pkg = "anthro"))
)
```

## `childsds`
```{r bench_childsds, eval = FALSE}

childsds_timings <- mbench_pkg(
  pkg_expr = quote(childsds::sds(value = y,
                                 age = x / 365.25,
                                 sex = sex,
                                 male = "M",
                                 female = "F",
                                 item = "weight",
                                 ref = childsds::who.ref)),
  pkg_name = paste("R: childsds", packageVersion(pkg = "childsds"))
)
```

## `growthstandards`
```{r bench_growthstandards, eval = FALSE}
growthstandards_timings <- mbench_pkg(
  pkg_expr = quote(
    growthstandards::who_value2zscore(y = y, x = x,
                                      sex = ifelse(sex == "M", "Male",
                                                   "Female"),
                                      y_var = "wtkg")
  ),
  pkg_name = paste("R: growthstandards",
                   packageVersion(pkg = "growthstandards"))
)
```

## `gigs`
```{r bench_gigs, eval = FALSE}
gigs_timings <- mbench_pkg(
  pkg_expr = quote(gigs::who_gs_value2zscore(y = y, x = x, sex = sex,
                                             acronym = acronym)),
  pkg_name = paste("R: gigs", packageVersion(pkg = "gigs"))
)
```

## Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(data = bench_dataset,
                 path = file.path("exclude", "statabench", "bench_dataset.dta"))
```

In Stata, the commands are run inside a do-file which utilises the
[benchmark](https://www.github.com/mcaceresb/stata-benchmark) package for Stata.
This script essentially does the same as `mbench_pkg()`, but for the
[Stata gigs package](https://www.github.com/lshtm-gigs/gigs-stata/) and the
zanthro package.
```{stata stata_bench_code, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_dataset.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = who_gs(y, "wfa", "v2z"), ///
		xvar(x) sex(sex) sexcode(m=M, f=F)
}

foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_dataset.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen z_anthro = zanthro(y, wa, WHO), xvar(x) gender(sex) ///
                gencode(male=M, female=F)  ageunit(day)
}
```

```{r bind_stata_timings, echo = FALSE, eval = FALSE}
possible_n <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_3_0 <- c(0.039, 0.039, 0.040, 0.041, 0.042, 0.057, 0.076, 0.134, 0.230,
                0.332, 0.440)
zbmicat_1_0_2 <- c(0.007, 0.008, 0.009, 0.017, 0.027, 0.105, 0.203, 0.479,
                   0.951, 1.4450, 2.046)
lens <- lengths(list(gigs_0_3_0, zbmicat_1_0_2))
n_inputs <- unlist(lapply(X = lens, FUN = \(x) possible_n[seq_len(x)]))
stata_timings <-  data.frame(n_inputs = n_inputs,
                             median_time = c(gigs_0_3_0, zbmicat_1_0_2),
                             time_units = "seconds",
                             package = c(rep("Stata: gigs 0.3.0", lens[1]),
                                         rep("Stata: zbmicat 1.0.2", lens[2])))
rm(gigs_0_3_0, zbmicat_1_0_2, lens, n_inputs, possible_n)
```

The outputs from this script give a table of timings that look like this:
```{r show_stata}
stata_timings
```

## Plotting
Only by plotting the various timings tables can we see the trends for each
package:

```{r plot_timings, echo = FALSE}
dplyr::bind_rows(anthro_timings,
                 childsds_timings,
                 growthstandards_timings,
                 gigs_timings,
                 stata_timings) |>
  dplyr::mutate(median_ms = dplyr::case_when(
    time_units == "nanoseconds" ~ median_time / 1000000,
    time_units == "microseconds" ~ median_time / 1000,
    time_units == "milliseconds" ~ median_time,
    time_units == "seconds" ~ median_time  * 1000.,
  )) |>
  ggplot2::ggplot(ggplot2::aes(x = n_inputs,
                               y = median_ms,
                               colour = package)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(colour = "black") +
  ggplot2::labs(x = "Number of inputs",
                y = "Median time taken (milliseconds)",
                colour = "Package") +
  ggsci::scale_colour_lancet()
```

On the whole, `anthro` is by far the slowest R package, taking around 2.56
seconds to run over 100,000 inputs. This is in part because `anthro` computes
results in every WHO Child Growth standard each time anthro is called, but also
due to a slower implementation than the other packages.

Next slowest is the Stata package `zanthro`, which takes 2.05 seconds to compute
results in just one WHO standard. About 4 times faster than `zanthro` is `gigs`
for Stata, which scales more efficiently than `zanthro` and so takes 0.44
seconds to convert 100,000 measurements to z-scores.

Leading the pack are three R implementations: `growthstandards`, `gigs`, and
`childsds`. The `childsds` package is fastest at ~ 145 ms for 100,000 inputs,
followed by `growthstandards` (166 ms) and `gigs` (168 ms).

# Package output similarity
The packages also differ slightly in how they convert between different values,
which can affect the z-scores computed by each package, even if inputs are the
same. In our testing, we found that the tested packages mostly agreed with each
other, but that `childsds` did not correctly perform z-scoring the WHO
standards.

This is because the WHO Child Growth standards constrain z-scores in the outer
tails to within the z-scores where more data was available, i.e. between -3 and
+3 SD. More information on this can be found in the reports referenced in the
`gigs::who_gs_value2zscore()` documentation.

```{r discrepancies, eval = FALSE}
discrepancies <- data.frame(z = c(-3.03, -2.97, 2.97, 3.03),
                            age_days = 0,
                            sex = "M") |>
  dplyr::mutate(
    weight_kg = gigs::who_gs_wfa_zscore2value(z, age_days, sex),
    # GIGS z-score
    z_gigs = gigs::who_gs_wfa_value2zscore(weight_kg, age_days, sex),
    z_growthstandards = growthstandards::who_wtkg2zscore(
      age_days, weight_kg, "Male"
    ),
    z_childsds = childsds::sds(
      value = weight_kg, age = age_days / 365.25,
      sex = sex, male = "M", female = "F",
      item = "weight", ref = childsds::who.ref
    )
  )
```

When we look at these z-scores, you can see that both `growthstandards` and
`gigs` correctly apply the constraining procedure; `childsds` does not.

```{r discrepancies_kable, echo = FALSE}
knitr::kable(discrepancies, align = "ccccc")
```

# Session information
```{r session_info1, echo = TRUE, eval = FALSE}
sessionInfo()
```

```{r session_info2, echo = FALSE, eval = FALSE}
session_info <- sessionInfo()
```

```{r session_info3, echo = FALSE, eval = TRUE}
session_info
```



[//]: # (Code to save/load all timings tables in a separate .rda file)
```{r save_benchmarking_rda, eval = FALSE, echo = FALSE}
save(file = file.path("vignettes/articles/benchmarking.rda"),
     anthro_timings, childsds_timings, growthstandards_timings, gigs_timings,
     stata_timings, discrepancies, session_info)
```