---
title: "Benchmarking gigs against other software packages"
description: >
  A comparison of gigs' performance against other growth analysis packages.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dpi = 120, message = FALSE, warning = FALSE,
  dev.args = list(png = list(type = "cairo"))
)

# Set ggplot theming for consistent plots
ggplot2::theme_set(ggplot2::theme_bw())

intergrowth21st <- if (knitr::is_html_output()) {
  knitr::asis_output(x = "INTERGROWTH-21<sup>st</sup>")
} else {
  knitr::asis_output(x = "INTERGROWTH-21\textsupercript{st}")
}
```

```{r srr-tags, eval = FALSE, echo = FALSE}
#' @srrstats {1.5, 1.6} This article contains code for both performance and
#'   accuracy claims when comparing this package with other software packages.
```

```{r load_timings_rda, echo = FALSE}
load(file = file.path("benchmarking.rda"))
```

# Rationale
This short article describes the performance of gigs relative to a
non-exhaustive group of R and non-R packages which implement international
growth standards which are also found in gigs.

```{r pkgnames, echo = FALSE}
anthro <- knitr::asis_output(x = "[anthro](https://cran.r-project.org/web/packages/anthro/index.html)")
AGD <- knitr::asis_output(x = "[AGD](https://cran.r-project.org/web/packages/AGD/index.html)")
childsds <- knitr::asis_output(x = "[childsds](https://cran.r-project.org/web/packages/childsds/index.html)")
gigs_r <- knitr::asis_output(x = "[gigs](https://www.github.com/lshtm-gigs/gigs/)")
growthstans <- knitr::asis_output(x = "[ki-tools/growthstandards](https://www.github.com/ki-tools/growthstandards/)")
sitar <- knitr::asis_output(x = "[sitar](https://cran.r-project.org/web/packages/sitar/index.html)")
zscorer <- knitr::asis_output(x = "[zscorer](https://cran.r-project.org/web/packages/zscorer/index.html)")
gigs_stata <- knitr::asis_output(x = "[gigs](https://www.github.com/lshtm-gigs/gigs-stata/) (Stata)")
zanthro <- knitr::asis_output(x = "[zanthro](https://journals.sagepub.com/doi/epdf/10.1177/1536867X1301300211) (Stata)")
intergrowth <- knitr::asis_output(x = "[nutriverse/intergrowth](https://github.com/nutriverse/intergrowth/)")
```

| Package name    | Language | On CRAN? |
|-----------------|:---------|:---------|
| `r gigs_r`      | R        | Yes      |
| `r anthro`      | R        | Yes      |
| `r AGD`         | R        | Yes      |
| `r childsds`    | R        | Yes      |
| `r sitar`       | R        | Yes      |
| `r zscorer`     | R        | Yes      |
| `r growthstans` | R        | No       |
| `r intergrowth` | R        | No       |
| `r gigs_stata`  | Stata    | No       |
| `r zanthro`     | Stata    | No       |

Thus far, there is no comprehensive benchmark comparing these different
packages. This short article will compare the speed of each package from 1 to
100000 inputs, checking how fast each package can convert weight values to
z-scores in different growth standards.

We performed these benchmarks on R version
`r session_info[[1]][["major"]]`.`r session_info[[1]][["minor"]]`, using a
Windows 10 system with a Ryzen 7 3700X processor and 16GB of DDR4 RAM. The Stata
benchmarks were run in Stata 18.0 (revision 15 May 2023) on the same system,
using the [benchmark](https://github.com/mcaceresb/stata-benchmark) package for
Stata.

```{r yes_no_half, echo = FALSE}
yes <- if (knitr::is_html_output()) {
   knitr::asis_output(x = "✅")
} else {
  knitr::asis_output(x = "✓")
}
no <- if (knitr::is_html_output()) {
  knitr::asis_output(x = "❌")
} else {
  knitr::asis_output(x = "✘")
}
half <- if (knitr::is_html_output()) {
  knitr::asis_output(x = "⚠️")
} else {
  knitr::asis_output(x = "⚠")
}
```

Other R packages can be used to analyse growth data with international
standards, but have limitations which are not present in gigs. There are also
software packages external to R which implement these standards. The table
below describes these packages, and to what extent they have implemented
functions that let users convert anthropometric measurements to
z-scores/centiles in each set of standards implemented in gigs - the WHO Child
Growth standards, `r intergrowth21st` Newborn Size standards (including Very
Preterm), and the `r intergrowth21st` Postnatal Growth standards for preterm
infants. A tick (`r yes`) indicates that a package contains functions for all
growth standards within a set, a cross (`r no`) indicates that a package has no
functions for a set of growth standards, and a warning sign (`r half`) indicates
that some standards in a set are implemented, but not others.

```{r arrows, echo = FALSE}
ig21st <- if (knitr::is_html_output()) {
  knitr::asis_output(x = "IG-21<sup>st</sup>")
} else {
  knitr::asis_output(x = "IG-21\textsupercript{st}")
}

larr <- knitr::asis_output(x = "&larr;")
rarr <- knitr::asis_output(x = "&rarr;")
harr <- knitr::asis_output(x = "&harr;")
```

| Software        | Platform | WHO (0-5 years) | `r ig21st` NBS | `r ig21st` PNG | `r ig21st` Fetal | Functionality                     |
|-----------------|----------|-----------------|----------------|----------------|------------------|-----------------------------------|
| `r gigs_r`      | R        | `r yes`         | `r yes`        | `r yes`        | `r yes`          | Values `r harr` z-scores/centiles |
| `r anthro`      | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores          |
| `r AGD`         | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r harr` z-scores          |
| `r childsds`    | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores/centiles |
| `r growthstans` | R        | `r yes`         | `r yes`        | `r half`       | `r half`         | Values `r harr` z-scores/centiles |
| `r intergrowth` | R        | `r no`          | `r no`         | `r no`         | `r half`         | Values `r rarr` z-scores/centiles |
| `r sitar`       | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r harr` z-scores/centiles |
| `r zscorer`     | R        | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores/centiles |
| `r gigs_stata`  | Stata    | `r yes`         | `r yes`        | `r yes`        | `r yes`          | Values `r harr` z-scores/centiles |
| `r zanthro`     | Stata    | `r yes`         | `r no`         | `r no`         | `r no`           | Values `r rarr` z-scores/centiles |

# Shared benchmark code

```{r load_gigs}
library(gigs)
```

In this vignette, we perform benchmark performance in each a growth standard
from each set of growth standards implemented in gigs. These are the:

* WHO child growth standards (0-5 years)
* `r intergrowth21st` Newborn Size standards
* `r intergrowth21st` Postnatal Growth standards
* `r intergrowth21st` Fetal standards

The benchmarks utilise pre-generated datasets for growth standard. The z-scores,
`x` variable and sexes for each row are generated (pseudo)randomly. Because the
`gigs` test suite validates its functions against published growth curve data
from the WHO and the `r intergrowth21st` Project, we use `gigs` to generate
expected measurements for each observation. This dataset can then be used to
compare speeds for each package. These datasets are generated with the following
function:
```{r benchmark_dataset, echo = TRUE}
benchmark_dataset <- function(n, family, acronym) {
  data <- get(family, envir = asNamespace("gigs"))
  if (family == "ig_fet") {
    xvars <- data[[acronym]][[1]][[1]]
  } else {
    xvars <- data[[acronym]][[1]][[1]][[1]]
  }

  withr::with_seed(seed = 154237890, code = {
      # Random z-scores around 0
      z <- rnorm(n = n)
    })
  withr::with_seed(seed = 154237890, code = {
      # X variables are non-discrete but within bounds of the desired acronym
      x <-  sample(xvars, size = n, replace = TRUE)
      x_jitter <- c(runif(n = 5, min = -1, max = 1), 0)
      x <- x + sample(x_jitter, size = n, replace = TRUE)
      x <- pmax(pmin(x, max(xvars)), min(xvars))
      # Sexes randomly sampled from male and female
      sex <- sample(c("M", "F"), size = n, replace = TRUE)
    }
  )
  gigs_fn <- get("zscore2value", envir = asNamespace("gigs"))
  out <- if (family != "ig_fet") {
    list(z = z, x = x, sex = sex, family = family, acronym = acronym)
  } else {
    list(z = z, x = x, family = family, acronym = acronym)
  }
  out$y <- do.call(what = gigs_fn, args = out)
  if (family == "ig_fet") {
    out$sex <- NA
  }
  as.data.frame(out, stringsAsFactors = FALSE)
}
```

```{r hidden_helpers, echo = FALSE}
max_median <- function(bp, name) {
  round(max(bp[bp$desc == name, "median"]))
}

pkg_str <- function(pkgname) paste("R:", pkgname, packageVersion(pkg = pkgname))

ver_stata_gigs <- "0.4.0"
ver_stata_zanthro <- "1.0.2"
plot_bp <- function(bp_r, bp_stata = NULL) {
  key_pkg_colour <- c("gs" = "#925E9FFF",
                      "gigs" = "#00468BFF",
                      "anthro" = "#AD002AFF",
                      "AGD" = "cornflowerblue",
                      "intergrowth" = "#FDAF91FF",
                      "childsds" = "#663300FF",
                      "sitar" = "#005500FF",
                      "zscorer" = "#009900",
                      "stata_gigs" = "skyblue",
                      "stata_zanthro" = "darkolivegreen4")
  key_pkg_label <- c("gs" = pkg_str("growthstandards"),
                     "gigs" = pkg_str("gigs"),
                     "anthro" = pkg_str("anthro"),
                     "AGD" = pkg_str("AGD"),
                     "intergrowth" = pkg_str("intergrowth"),
                     "childsds" = pkg_str("childsds"),
                     "sitar" = pkg_str("sitar"),
                     "zscorer" = pkg_str("zscorer"),
                     "stata_gigs" = paste("Stata: gigs", ver_stata_gigs),
                     "stata_zanthro" = paste("Stata: zanthro",
                                             ver_stata_zanthro))
  bp <- bp_r  |>
    dplyr::select(desc, input_len, median)
  if (!is.null(bp_stata)) {
    suppressMessages(suppressWarnings(bp <- dplyr::full_join(bp, bp_stata)))
  }
  bp |>
    ggplot2::ggplot(ggplot2::aes(x = input_len, y = median, colour = desc)) +
    ggplot2::geom_line() +
    ggplot2::geom_point(colour = "black") +
    ggplot2::labs(x = "Input lengths", y = "Median time taken (ms)",
                  colour = "Package") +
    ggplot2::scale_colour_manual(values = key_pkg_colour,
                                 labels = key_pkg_label) +
    ggplot2::theme_bw()
}
```

We're using the `{bench}` package throughout this vignette, specifically the
`bench::press()` function, which lets us compare function scaling by iterating
through increasing amounts of observations.

# WHO Growth Standards
The World Health Organisation (WHO) Child Growth Standards are included in gigs
with the various `who_gs_*`conversion functions. These standards are implemented
in a variety of other packages.

## Set up benchmark dataset

```{r ds_who_gs}
# Generate 100,000-row dataset
ac_who_gs <- "wfa"
ds_who_gs <- benchmark_dataset(n = 100000,
                               family = "who_gs",
                               acronym = ac_who_gs)
```

The first 10 rows of this dataset look like this:
```{r show_bench_dataset_who_gs, echo = FALSE}
ds_who_gs[1:10, ]
```


## Timing
### R
```{r bp_who_gs, echo = TRUE, eval = FALSE}
temp_bp_who_gs_most <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    selector <- seq_len(input_len)
    weight_kg <- ds_who_gs$y[selector]
    age_days <- ds_who_gs$x[selector]
    acronym <- ds_who_gs$acronym[selector]
    sex <- ds_who_gs$sex[selector]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    age_years <- age_days / 365.25
    bm_who_gs <- bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 30,
      gigs = gigs::value2zscore(weight_kg, age_days, sex,
                                family = "who_gs", acronym = ac_who_gs),
      anthro = anthro::anthro_zscores(weight = weight_kg,
                                      age = age_days,
                                      sex = sex)$zwei,
      AGD = AGD::y2z(y = weight_kg,
                     x = age_days,
                     sex = sex,
                     ref = AGD::who.wgt,
                     tail.adjust = TRUE),
      childsds = childsds::sds(value = weight_kg,
                              age = age_years,
                              sex = sex,
                              male = "M",
                              female = "F",
                              item = "weight",
                              ref = childsds::who.ref),
      gs = growthstandards::who_value2zscore(y = weight_kg, x = age_days,
                                             y_var = "wtkg", x_var = "agedays",
                                             sex = sex_gs),
      sitar = sitar::LMS2z(y = weight_kg, x = age_years, sex = sex,
                          measure = "wt", ref = sitar::who06)
    )
  })

temp_bp_who_gs_zscorer <- bench::press(
  input_len = c(1, 1000, 5000, 10000, 15000),
  {
    selector <- seq_len(input_len)
    weight_kg <- ds_who_gs$y[selector]
    age_days <- ds_who_gs$x[selector]
    sex <- ds_who_gs$sex[selector]
    df_zscorer <- data.frame(
      weight = weight_kg, age = age_days,
      sex = ifelse(sex == "M", 1, 2)
    )
    bm_who_gs <- bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", max_iterations = 5,
      zscorer = zscorer::addWGSR(
        df_zscorer, sex = "sex", firstPart = "weight", secondPart = "age",
        index = "wfa"
      )
    )
  })

bp_who_gs <- dplyr::bind_rows(temp_bp_who_gs_most, temp_bp_who_gs_zscorer) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_who_gs,
  path = file.path("exclude", "statabench", "bench_ds_who_gs.dta")
)
```

In Stata, the commands are run inside a do-file which utilises the
[benchmark](https://www.github.com/mcaceresb/stata-benchmark) package for Stata. This code essentially does the same as
`bench::press()`, but for the packages we're testing in Stata.
```{stata stata_bench_code_who_gs, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_who_gs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = who_gs(y, "wfa", "v2z"), ///
		    xvar(x) sex(sex) sexcode(m=M, f=F)
}

foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_who_gs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen z_anthro = zanthro(y, wa, WHO), xvar(x) gender(sex) ///
            gencode(male=M, female=F) ageunit(day)
}
```

```{r bp_who_gs_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.009, 0.008, 0.009, 0.010, 0.012, 0.028, 0.047, 0.104, 0.198,
                0.296, 0.405)
zanthro_1_0_2 <- c(0.007, 0.008, 0.009, 0.017, 0.027, 0.105, 0.203, 0.479,
                   0.951, 1.4450, 2.046)
lens <- lengths(list(gigs_0_4_0, zanthro_1_0_2))
n_inputs <- unlist(lapply(X = lens, FUN = \(x) input_len[seq_len(x)]))
bp_who_gs_stata <-  data.frame(input_len = n_inputs,
                               median = c(gigs_0_4_0, zanthro_1_0_2) * 1000,
                               desc = c(rep("stata_gigs", lens[1]),
                                        rep("stata_zanthro", lens[2])))
rm(gigs_0_4_0, zanthro_1_0_2, lens, n_inputs, input_len)
```

## Package comparisons: timing

```{r plot_timings_who_gs, echo = FALSE, fig.alt = "A line graph of median time taken against no. of inputs for different software packages. The results can be read about below."}
plot_bp(bp_who_gs, bp_who_gs_stata)
```

On the whole, `zscorer` is by far the slowest R package, taking around
`r max_median(bp_who_gs, "zscorer") / 1000` seconds to process 100,000
inputs. Looking at the source code, this is because `zscorer::addWGSR()` hasn't
been vectorised. Next slowest is `anthro`, which takes
`r max_median(bp_who_gs, "anthro") / 1000` seconds to process 100,000
inputs. This slowdown arises from `anthro` computing results in every WHO Child
Growth standard each time `anthro::anthro_zscores()` is called, but also due to
a slower implementation of the WHO LMS procedure than the other R packages.

Next slowest is the Stata package `zanthro`, which takes around
`r max_median(bp_who_gs_stata, "stata_zanthro") / 1000` seconds to compute
results in a single WHO standard. About 4 times faster than `zanthro` is `gigs`
for Stata, which scales more efficiently than `zanthro` and takes around
`r max_median(bp_who_gs_stata, "stata_gigs") / 1000` seconds to convert 100,000
measurements to z-scores.

Focussing on the faster R implementations reveals some interesting patterns:

```{r plot_timings_who_gs_faster, echo = FALSE, fig.alt = "A line graph of median time taken against no. of inputs for a faster subset of the software packages analysed. The results can be read about below."}
bp_who_gs |>
  dplyr::filter(desc %in% c("gigs", "childsds", "gs", "sitar", "AGD")) |>
  plot_bp()
```

The `sitar` package is far and away the fastest here, taking
~`r max_median(bp_who_gs, "sitar")` ms to process 100,000 inputs. This is because
it uses the monthly LMS coefficients to calculate its z-scores. Whilst this is
quicker, it does induce some imprecision when compared to the other packages.

Next fastest was the `growthstandards` package at
~`r max_median(bp_who_gs, "gs")` ms for 100,000 inputs, followed by `gigs`
(~`r max_median(bp_who_gs, "gigs")` ms), `childsds`
(~`r max_median(bp_who_gs, "childsds")` ms), and lastly `AGD`
(~`r max_median(bp_who_gs, "AGD")` ms). Interestingly, `AGD` starts out
much slower than the other 'fast' packages, but may scale more efficiently.

## Package comparisons: numerical consistency
In our testing of the WHO standards, we found that the tested packages mostly
agreed with each other when provided the same inputs, except for `childsds` and
`sitar`.

For `childsds`, this is because the WHO Child Growth standards constrain
z-scores in the outer tails (i.e. past 3 z-scores), as their data was more
sparse for these extreme values. More information on this constraining procedure
can be found in the reports referenced in the `gigs::who_gs_value2zscore()`
documentation. It appears that `childsds` does not perform this constraining
procedure, so conversions where `abs(z_score) > 3` are computed incorrectly:
```{r discrepancies_who_gs}
discrepancies <- data.frame(z = c(-3.03, -2.97, 2.97, 3.03),
                            age_days = 0,
                            sex = "M") |>
  dplyr::mutate(
    weight_kg = gigs::zscore2value(z, age_days, sex, "who_gs", "wfa"),
    # GIGS z-score
    z_gigs = gigs::value2zscore(weight_kg, age_days, sex, "who_gs", "wfa"),
    # growthstandards z-score
    z_growthstandards = growthstandards::who_wtkg2zscore(
      age_days, weight_kg, "Male"
    ),
    # childsds z-score
    z_childsds = childsds::sds(
      value = weight_kg, age = age_days / 365.25,
      sex = sex, male = "M", female = "F",
      item = "weight", ref = childsds::who.ref
    )
  )
```

When we look at these z-scores, you can see that both `growthstandards` and
`gigs` correctly apply the constraining procedure; `childsds` does not. The
`AGD` package constrains z-scores if the `tail.adjust` option in `AGD::y2z()` is
specified, and from looking at the `anthro` source code, they also apply the
constraining procedure.
```{r discrepancies_who_gs_kable, echo = FALSE}
knitr::kable(discrepancies, align = "ccccc")
```

Discrepancies between `sitar` and `gigs` arise from `sitar` relying on monthly
LMS values for the WHO Growth Standards, instead of the daily tables. This means
for the same day, `sitar` and `gigs` will use slightly different lambda/mu/sigma
values, and so compute different z-scores for the same measurements.

# INTERGROWTH-21<sup>st</sup> Newborn Size standards
The `r intergrowth21st` Newborn Size standards are implemented in gigs for R and
Stata, and in the `growthstandards` package for R. Let's make a new dataset,
`ds_ig_nbs`, which we will use to benchmark these functions:

```{r ds_ig_nbs}
ac_ig_nbs <- "wfga"
ds_ig_nbs <- benchmark_dataset(n = 100000,
                               family = "ig_nbs",
                               acronym = ac_ig_nbs)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset_ig_nbs}
ds_ig_nbs[1:10, ]
```

## Timing
### R
```{r bp_ig_nbs, echo = TRUE, eval = FALSE}
bp_ig_nbs <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    p <- pnorm(ds_ig_nbs$z[1:input_len])
    weight_kg <- ds_ig_nbs$y[1:input_len]
    gest_days <- ds_ig_nbs$x[1:input_len]
    acronym <- ds_ig_nbs$acronym[1:input_len]
    sex <- ds_ig_nbs$sex[1:input_len]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::value2centile(y =weight_kg, x = gest_days, sex = sex,
                                 family = "ig_nbs", acronym = ac_ig_nbs),
      gs = growthstandards::igb_value2centile(val = weight_kg,
                                              gagebrth = gest_days,
                                              sex = sex_gs, var = "wtkg")
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```
### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_nbs,
  path = file.path("exclude", "statabench", "bench_ds_ig_nbs.dta")
)
```

We can then benchmark the speed of the Stata command:
```{stata stata_bench_code_ig_nbs, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_nbs.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_nbs(y, "wfga", "v2z"), ///
		    gest_days(x) sex(sex) sexcode(m=M, f=F)
}
```

```{r bp_ig_nbs_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.009, 0.004, 0.005, 0.006, 0.008, 0.027, 0.048, 0.111, 0.220,
                0.334, 0.471)
bp_ig_nbs_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_nbs, echo = FALSE, fig.alt = "A line graph of median time taken against no. of inputs for the analysed software packages in the INTERGROWTH-21st Newborn Size standard for weight-for-gestational age. The results can be read about below."}
plot_bp(bp_ig_nbs, bp_ig_nbs_stata)
```

For this set of growth standards, the Stata implementation for GIGS is the
slowest at ~`r max_median(bp_ig_nbs_stata, "stata_gigs")` ms. The
`growthstandards` and `gigs` packages are the fastest at
~`r max_median(bp_ig_nbs, "gs")` and ~`r max_median(bp_ig_nbs, "gigs")` ms,
respectively.

## Package comparisons: numerical consistency
In our testing of the `r intergrowth21st` Newborn Size standards, we found that
the implementation in `growthstandards` does not perform coefficient
interpolation for the `r intergrowth21st` Newborn Size standards which
utilise mu/sigma/nu/tau coefficients. Instead, `growthstandards` uses `round()`
to round non-integer gestational ages to the nearest value, then gets
mu/sigma/nu/tau coefficients for this rounded GA. This leads a smaller-scale
version of the z-scoring errors found by
[Kiger *et al.* (2016)](https://dx.doi.org/10.1007/s10916-015-0389-x) when not
interpolating LMS coefficients in the WHO standards.

```{r discrepancies_ig_nbs, eval = TRUE}
weight_kg <- 3
gestage <- seq(274, 278, 0.5)
gigs <- gigs::value2centile(y = weight_kg, x = gestage, sex = "M",
                            family = "ig_nbs", acronym = "wfga")
gs <- growthstandards::igb_value2centile(val = weight_kg, gagebrth = gestage,
                                         sex = "Male", var = "wtkg")
waldo::compare(gigs, gs / 100, x_arg = "gigs", y_arg = "growthstandards")
```

# INTERGROWTH-21<sup>st</sup> Postnatal Growth standards
The `r ig21st` Postnatal Growth standards are implemented in gigs for R and
Stata, and in the `growthstandards` package. Let's make a new dataset,
`ds_ig_png`, which we will use to benchmark these functions:
```{r ds_ig_png}
ac_ig_png <- "wfa"
ds_ig_png <- benchmark_dataset(n = 100000,
                               family = "ig_png",
                               acronym = ac_ig_png)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset4}
ds_ig_png[1:10, ]
```

## Timing
### R
```{r bp_ig_png, echo = TRUE, eval = FALSE}
bp_ig_png <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    weight_kg <- ds_ig_png$y[1:input_len]
    pma_weeks <- ds_ig_png$x[1:input_len]
    pma_days <- (pma_weeks * 7)[1:input_len]
    acronym <- ds_ig_png$acronym[1:input_len]
    sex <- ds_ig_png$sex[1:input_len]
    sex_gs <- ifelse(sex == "M", "Male", "Female")
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::value2zscore(weight_kg, pma_weeks, sex,
                                family = "ig_png", acronym = ac_ig_png),
      gs = growthstandards::igprepost_value2zscore(val = weight_kg,
                                                   pmagedays = pma_days,
                                                   sex = sex_gs, var = "wtkg")
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_png,
  path = file.path("exclude", "statabench", "bench_ds_ig_png.dta")
)
```

```{stata stata_bench_code_ig_png, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_png.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_png(y, "wfga", "v2z"), ///
		    xvar(x) sex(sex) sexcode(m=M, f=F)
}
```

```{r bp_ig_png_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.003, 0.002, 0.002, 0.003, 0.004, 0.010, 0.019, 0.043, 0.082,
                0.114, 0.164)
bp_ig_png_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_png, echo = FALSE, fig.alt = "A line graph of median time taken against no. of inputs for the analysed software packages in the INTERGROWTH-21st Postnatal Growth standard for weight-for-age. The results can be read about below."}
plot_bp(bp_ig_png, bp_ig_png_stata)
```

For the `r intergrowth21st` Postnatal Growth standards, the Stata implementation
for GIGS is the slowest at ~`r max_median(bp_ig_png_stata, "stata_gigs")` ms.
The `gigs` and `growthstandards` packages are the fastest at
~`r max_median(bp_ig_png, "gs")` and ~`r max_median(bp_ig_png, "gigs")` ms,
respectively.

# INTERGROWTH-21<sup>st</sup> Fetal standards
The `r ig21st` Fetal standards are implemented in gigs for R and Stata, and in
the `growthstandards` and `intergrowth` packages (though more fully in
`intergrowth`  than in `growthstandards`, and both are missing some standards
which are included in `gigs`). Let's make a new dataset, `ds_ig_fet`, which we
will use to benchmark a conversion in a fetal growth standard common to all
three packages:

```{r ds_ig_fet}
ac_ig_fet <- "ofdfga"
ds_ig_fet <- benchmark_dataset(n = 100000,
                               family = "ig_fet",
                               acronym = ac_ig_fet)
```


The first 10 rows of this dataset look like this:
```{r show_bench_dataset_ig_fet, eval = FALSE}
ds_ig_fet[1:10, ]
```

## Timing
### R
```{r bp_ig_fet, echo = TRUE, eval = FALSE}
bp_ig_fet <- bench::press(
  input_len = c(1, 1000, 10000, seq(25000, 100000, 25000)),
  {
    gest_days <- ds_ig_fet$x[1:input_len]
    gest_wks <- gest_days / 7
    acronym <- ds_ig_fet$acronym[1:input_len]
    ofd_mm <- ds_ig_fet$y[1:input_len]
    ofd_cm <- ofd_mm / 10
    bench::mark(
      check = FALSE, filter_gc = TRUE, time_unit = "ms", min_iterations = 50,
      gigs = gigs::value2zscore(ofd_mm, gest_days,
                                family = "ig_fet", acronym = ac_ig_fet),
      gs = growthstandards::igfet_value2zscore(val = ofd_cm,
                                               gagedays = gest_days,
                                               var = "ofdcm"),
      intergrowth = intergrowth::calculate_ofd_gestage(ofd = ofd_mm,
                                                       gestage = gest_wks)[[1]],
    )
  }) |>
  dplyr::mutate(desc = attr(expression, which = "description")) |>
  dplyr::select(!expression, !result, !memory, !time, !gc)
```

### Stata
```{r, eval = FALSE}
# Save .dta file equivalent of benchmarking table. This can be used to benchmark
# Stata packages.
haven::write_dta(
  data = ds_ig_fet,
  path = file.path("exclude", "statabench", "bench_ds_ig_fet.dta")
)
```

```{stata stata_bench_code_ig_fet, eval = FALSE}
// This is Stata code
foreach i in 1 10 100 500 1000 5000 10000 25000 50000 75000 100000 {
	use "benchmarking/bench_ds_ig_fet.dta", clear
	qui drop if _n > `i'
	di "Number of inputs: `i'"
	bench, reps(25) restore last: ///
		qui egen double z_gigs = ig_fet(y, "ofdfga", "v2z"), xvar(x)
}
```

```{r bp_ig_fet_stata, echo = FALSE, eval = FALSE}
input_len <- c(1, 10, 100, 500, 1000, 5000, 10000, seq(25000, 100000, 25000))
gigs_0_4_0 <- c(0.002, 0.003, 0.002, 0.003, 0.004, 0.006, 0.012, 0.025, 0.046,
                0.066, 0.093)
bp_ig_fet_stata <-  data.frame(input_len = input_len,
                               median = gigs_0_4_0 * 1000,
                               desc = "stata_gigs")
rm(gigs_0_4_0, input_len)
```

```{r plot_timings_ig_fet, echo = FALSE, fig.alt = "A line graph of median time taken against no. of inputs for the analysed software packages in the INTERGROWTH-21st Fetal standard for occipito-frontal diameter-for-gestational age. The results can be read about below."}
plot_bp(bp_ig_fet, bp_ig_fet_stata)
```

For the `r intergrowth21st` Fetal standard for occipito-frontal diameter, the
GIGS implementation for Stata is the slowest to analyse 100,000 inputs at
~`r max_median(bp_ig_fet_stata, "stata_gigs")` ms. The R `gigs` package is the
fastest at ~`r max_median(bp_ig_fet, "gigs")` ms, followed by `growthstandards`
at ~`r max_median(bp_ig_fet, "gs")`, and then `intergrowth` at
~`r max_median(bp_ig_fet, "intergrowth")`.

# Summary

## Timings summary (100,000 inputs)

```{r save_timings, echo = FALSE, eval = FALSE}
max_median_all <- function(bp_r, bp_stata) {
  bp_both <- dplyr::bind_rows(bp_r[, colnames(bp_stata)], bp_stata)
  vapply(X = unique(bp_both$desc),
         FUN = \(desc) max_median(bp_both, desc),
         FUN.VALUE = double(1L))
}

max_len_bench <- list(
  who_gs = max_median_all(bp_who_gs, bp_who_gs_stata),
  ig_nbs = max_median_all(bp_ig_nbs, bp_ig_nbs_stata),
  ig_png = max_median_all(bp_ig_png, bp_ig_png_stata),
  ig_fet = max_median_all(bp_ig_fet, bp_ig_fet_stata)
)
```

| Software        | Platform | WHO (0-5 years) (ms)                      | `r ig21st` NBS (ms)                    | `r ig21st` PNG (ms)                    | `r ig21st` Fetal (ms)                   |
|-----------------|----------|-------------------------------------------|----------------------------------------|----------------------------------------|-----------------------------------------|
| `r gigs_r`      | R        | `r max_len_bench[[1]][["gigs"]]`          | `r max_len_bench[[2]][["gigs"]]`       | `r max_len_bench[[3]][["gigs"]]`       | `r max_len_bench[[4]][["gigs"]]`        |
| `r anthro`      | R        | `r max_len_bench[[1]][["anthro"]]`        | `r no`                                 | `r no`                                 | `r no`                                  |
| `r AGD`         | R        | `r max_len_bench[[1]][["AGD"]]`           | `r no`                                 | `r no`                                 | `r no`                                  |
| `r childsds`    | R        | `r max_len_bench[[1]][["childsds"]]`      | `r no`                                 | `r no`                                 | `r no`                                  |
| `r growthstans` | R        | `r max_len_bench[[1]][["gs"]]`            | `r  max_len_bench[[2]][["gs"]]`        | `r max_len_bench[[3]][["gs"]]`         | `r max_len_bench[[4]][["gs"]]`          |
| `r intergrowth` | R        | `r no`                                    | `r no`                                 | `r no`                                 | `r max_len_bench[[4]][["intergrowth"]]` |
| `r sitar`       | R        | `r max_len_bench[[1]][["sitar"]]`         | `r no`                                 | `r no`                                 | `r no`                                  |
| `r zscorer`     | R        | NA                                        | `r no`                                 | `r no`                                 | `r no`                                  |
| `r gigs_stata`  | Stata    | `r max_len_bench[[1]][["stata_gigs"]]`    | `r max_len_bench[[2]][["stata_gigs"]]` | `r max_len_bench[[3]][["stata_gigs"]]` | `r max_len_bench[[4]][["stata_gigs"]]`  |
| `r zanthro`     | Stata    | `r max_len_bench[[1]][["stata_zanthro"]]` | `r no`                                 | `r no`                                 | `r no`                                  |

Note: `zscorer` is NA because we couldn't time it for 100,000 inputs (it takes
too long).

## Why GIGS?
Overall, though the `gigs` package for R isn't always the fastest, it is only
some milliseconds off, and offers:

  * A full complement of international growth standards (i.e. no
    half-implemented sets of growth standards)
  * More accurate implementations in some cases
  * Growth categorisation functions

Some users may want to apply growth references in their work - that's fine, and
other R packages (e.g. `childsds` in R and `zanthro` in Stata) offer reference
charts which may be of use to end-users. However, for international growth
standards, we believe we offer the best combination of speed and range of
standards.

In R, `gigs` also offers input checking via its options, and will tell you if
you're providing non-workable data. We think this outweighs a couple of
milliseconds' slowdown. In Stata, where `gigs` consistently runs faster and
offers more growth standards than `zanthro`, `gigs` is the obvious choice.

# Session information
```{r session_info1, echo = TRUE, eval = FALSE}
sessionInfo()
```

```{r session_info2, echo = FALSE, eval = FALSE}
session_info <- sessionInfo()
```

```{r session_info3, echo = FALSE, eval = TRUE}
session_info
```



[//]: # (Code to save/load all timings tables in a separate .rda file)
```{r save_benchmarking_rda, eval = FALSE, echo = FALSE}
save(file = file.path("vignettes/articles/benchmarking.rda"),
     list = ls(pattern = "^(bp)"), max_len_bench, session_info)
```