---
title: "Extending the INTERGROWTH-21st Newborn Size Standards"
description: >
  Our process for extending the GA range INTERGROWTH-21st Newborn Size
  standards.
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(comment = "#>",
                      dev.args = list(png = list(type = "cairo")))

# Set ggplot theming for consistent plots
ggplot2::theme_set(ggplot2::theme_bw())

intergrowth21st <- if (knitr::is_html_output()) {
  knitr::asis_output(x = "INTERGROWTH-21<sup>st</sup>")
} else {
  knitr::asis_output(x = "INTERGROWTH-21\textsupercript{st}")
}
```

# Setup

```{r load_gigs}
library(gigs)
library(mfp)
```

# Rationale
Over the years, there have been many requests from users of the
`r intergrowth21st` Newborn Size Standards (including the Very Preterm
standards) to extend their range from 168 to 300 days (24<sup>+0</sup> to
42<sup>+6</sup> weeks) gestational age (GA) to a wider span of 154 to 314 days
(22<sup>+0</sup> to 44<sup>+6</sup> weeks) GA. We've decided to implement this
in **gigs** as an entirely separate standard from the `r intergrowth21st`
Newborn Size standards, to make these extended versions of the standards
available to all.

We want to note here that this process has been checked over by the former Lead
Statistician for the `r intergrowth21st` Project, Eric Ohuma, but is not
endorsed officially by `r intergrowth21st`. We want users to be aware that
z-scores/centiles derived from these extended bounds must be used with caution -
it is at the user's discretion to ensure they are careful with these extended
standards.

# Approach
## For the Very Preterm standards
For the `r intergrowth21st` Very Preterm standards, extending the standards is
easy. The Very Preterm standards use simple equations to model the mean and
standard deviation at different gestational ages (GAs). Extending these
standards is as simple as using gestational age values from 154 to 167 in these
equations.

## For the Newborn Size standards
The `r intergrowth21st` Newborn Size standards are more difficult to extend as
they are coefficient-based. For each gestational age from 231 days
(33<sup>+0</sup> weeks) to 300 days (42<sup>+6</sup> weeks), there are
corresponding mu/sigma/nu/tau values. These are used with `gamlss.dist::pST3()`
to convert values to centiles in these standards, and with `gamlss.dist::qST3()`
to convert centiles to values. Each coefficient represents a different aspect of
the distribution of anthropometric measures at a given gestational age:

* `mu` - The median of the distribution at a given GA.
* `sigma` - The standard deviation of the distribution at a given GA.
* `nu` - The skewness of the distribution at a given GA.
* `tau` - The kurtosis of the distribution at a given GA.

To extend these standards out, we need to generate new coefficients for
gestational ages from 301 to 314 days (43<sup>+0</sup> to
44<sup>+6</sup> weeks). We will do this with fractional polynomial regression in
`{mfp}` package.

# Extrapolating coefficients for the Newborn Size Standards

The `mu` and `sigma` values are different across GAs for
each standard. Luckily, the `nu` and `tau` parameters are identical across all
gestational ages for most of these standards. Using `unique()`, we see that for
all `r intergrowth21st` Newborn Size standards except female length-for-GA
(`lfga`/`female`), `nu` and `tau` are constant.

```{r nu_tau_consistency, echo = FALSE}
for (acronym in c("wfga", "lfga", "hcfga")) {
  acro <- switch(acronym,
                 "wfga" = "Weight-for-GA",
                 "lfga" = "Length-for-GA",
                 "hcfga" = "Head circumference-for-GA")
  cat(paste0(acro, ":"))
  for (sex in c("male", "female")) {
    coeffs <- gigs::ig_nbs_coeffs[[acronym]][[sex]]
    cat(paste0("\n\t", switch(sex, "male" = "Male", "female" = "Female"), ":"))
    cat(paste("\n\t\t Unique `nu` values:", paste(unique(coeffs$nu),
                                                  collapse = ", ")))
    cat(paste("\n\t\t Unique `tau` values:", paste(unique(coeffs$tau),
                                                   collapse = ", ")))
  }
  if (acro != "hcfga") cat("\n\n")
}
```

Looking at the `nu` values for length-for-GA in females, we see that only the
first value is different (`0.9677560`). Every other value in this vector is
`0.9677568`:
```{r lfga_female_nu}
gigs::ig_nbs_coeffs$lfga$female$nu
```

We will assume for this set of coefficients that `nu` is equal to `0.967758`
after 300 days' GA; for all other sets of coefficients we know that `nu` and
`tau` are constant throughout.

The `mu` and `sigma` parameters for each standard require extrapolation. The
following function uses `mfp::mfp()` to model the slope of `mu` and  `sigma`
values against GA in each set of standards using fractional polynomials. The
resulting fractional polynomial produced by  can then be used with `predict()`
to extrapolate the slopes out to 314 days' GA.

```{r extrapolate_coeff}
extrapolate_msnt <- function(acronym, sex) {
  existing_coeffs <- gigs::ig_nbs_coeffs[[acronym]][[sex]]

  # Increase alpha cut-off from 0.05 to make `mfp()` 'take' to some standards
  alpha <- 0.15
  # Subset for fracpoly regresion; improves transition to >300 days vs. using
  # entire GA range
  modelling_cutoff <- 280
  modelling_coeffs <- dplyr::filter(existing_coeffs, gest_days >= modelling_cutoff)

  mfp_mu <- mfp::mfp(mu ~ fp(gest_days, df = 4), data = modelling_coeffs,
                     maxits = 200, alpha = alpha)

  coeffs_ext <- data.frame(gest_days = 231:314)
  coeffs_ext$mu <- predict(mfp_mu, newdata = coeffs_ext)

  mfp_sigma <- mfp::mfp(sigma ~ fp(gest_days, df = 4), data = modelling_coeffs,
                        maxits = 200, alpha = alpha)
  coeffs_ext$sigma <- predict(mfp_sigma, newdata = coeffs_ext)

  coeffs_ext$nu <- existing_coeffs$nu[2]
  coeffs_ext$tau <- existing_coeffs$tau[2]

  # Replace predicted coeffs with old coeffs for old GA range
  coeffs_ext$mu[seq_along(231:300)] <- existing_coeffs$mu
  coeffs_ext$sigma[seq_along(231:300)] <- existing_coeffs$sigma
  coeffs_ext
}
```

Let's iterate through our different standards and sexes. We store the results
in `ig_nbs_ext_coeffs`:

```{r extrapolate_wfga_male}
ig_nbs_ext_coeffs <- list(wfga = list(male = NULL, female = NULL),
                          lfga = list(male = NULL, female = NULL),
                          hcfga = list(male = NULL, female = NULL))

for (acronym in c("wfga", "lfga", "hcfga")) {
  for (sex in c("male", "female")) {
    ig_nbs_ext_coeffs[[acronym]][[sex]] <- extrapolate_msnt(acronym, sex)
  }
}
```

# Plotting the extrapolated coefficients

The dark grey dashed line marks the point at which we start using the original
coefficient values to fit a fractional polynomial to the data. The light grey
line shows where the extrapolation begins. Our plots show good extrapolations in
these coefficients.

```{r plot_extrapolated_msnt, echo = FALSE}
plot_extrapolated_msnt <- function(acronym, sex) {
  ext_coeffs <- ig_nbs_ext_coeffs[[acronym]][[sex]]

  plot_data <- data.frame(
    x = rep.int(ext_coeffs$gest_days, times = 2),
    y = c(ext_coeffs$mu, ext_coeffs$sigma),
    source = rep.int(c(rep_len("Original", 69),
                       rep_len("Extrapolated", 15)),
                     2),
    coeff = c(rep_len("mu", nrow(ext_coeffs)), rep_len("sigma", nrow(ext_coeffs)))
  )

  modelling_cutoff <- 280
  plot_data |>
    ggplot2::ggplot(ggplot2::aes(x = x, y = y, colour = source, group = 1)) +
    ggplot2::geom_line() +
    ggplot2::geom_vline(xintercept = 300, colour = "grey",
                        linetype = "dashed") +
    ggplot2::geom_vline(xintercept = modelling_cutoff, colour = "grey20",
                        linetype = "dashed") +
    ggplot2::scale_x_continuous(breaks = seq(231, 314, 14),
                                labels = seq(231, 314, 14)) +
    ggplot2::labs(
      x = "Gestational age (days)",
      y = "Coefficient value",
      colour = "Coefficient source") +
    ggplot2::facet_wrap(. ~ coeff, ncol = 2, scales = "free_y") +
    ggplot2::theme(legend.position = "bottom")
}
```

#### Weight-for-GA: male

```{r coeff_plot_wfga_male, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("wfga", "male")
```

#### Weight-for-GA: female

```{r coeff_plot_wfga_female, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis for weight-for-GA in males. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("wfga", "female")
```

#### Length-for-GA: male

```{r coeff_plot_lfga_male, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis for weight-for-GA in females. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("lfga", "male")
```

#### Length-for-GA: female

```{r coeff_plot_lfga_female, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis for length-for-GA in males. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("lfga", "female")
```

#### Head circumference-for-GA: male

```{r coeff_plot_hcfga_male, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis for head circumference-for-GA in males. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("hcfga", "male")
```

#### Head circumference-for-GA: female

```{r coeff_plot_hcfga_female, echo = FALSE, fig.alt = "Two line graphs showing coefficient value on the y-axis by gestational age (in days) on the x-axis for head circumference-for-GA in males. At 300 days, a vertical dashed grey line indicates that extrapolation of the values of the mu and sigma coefficients begins. After this line, the value of mu and sigma change similarly to before interpolation was carried out."}
plot_extrapolated_msnt("hcfga", "female")
```

# Plotting the extended standards

The following code has two functions. The first,
`generate_extrapolated_ig_nbs()`, produces a long data frame with anthropometry
for the 3<sup>rd</sup>, 10<sup>th</sup>, 25<sup>th</sup>, 50<sup>th</sup>,
75<sup>th</sup>, 90<sup>th</sup>, and 97<sup>th</sup> centiles. The second
function (`plot_extrapolated_ig_nbs()`) uses these to graph the centiles, with
grey dashed lines at 168 and 300 days to indicate where extrapolation of the
`r intergrowth21st` Very Preterm Newborn standards stops and extrapolation of
the extrapolation of the `r intergrowth21st` Newborn Size standards starts,
respectively. Each plot has a gap at 231 days (33<sup>+0</sup> weeks), where we
switch from using the `r intergrowth21st` Very Preterm standards to the
`r intergrowth21st` Newborn Size standards.

```{r hidden_extrapolation_fns}
generate_extrapolated_ig_nbs <- function(acronym, sex) {
  purrr::map2(
    .x = c(0.03, 0.1, 0.25, 0.5, 0.75, 0.9, 0.97),
    .y = c("P03", "P10", "P25", "P50", "P75", "P90", "P97"),
    .f = function(centile, name) {
      vpns_lim <- 231
      gest_days <- 154:314
      y_nbs_ext <- with(ig_nbs_ext_coeffs[[acronym]][[sex]],
                        gamlss.dist::qST3C(p = centile,
                                            mu = mu,
                                            sigma = sigma,
                                            nu = nu,
                                            tau = tau))
      y_vpns_ext <- gigs:::ig_vpns_zscore2value(z = qnorm(centile),
                                                154:230,
                                                switch(sex,
                                                       male = "M",
                                                       female = "F"),
                                                acronym)
      standard <- ifelse(
        test = gest_days >= vpns_lim,
        yes = "NS",
        no = "Very Preterm NS"
      )
      x <- data.frame(gest_days = gest_days,
                      Centile = name,
                      y = c(y_vpns_ext, y_nbs_ext),
                      standard = standard)
    }) |>
    purrr::list_rbind()
}

plot_extrapolated_chart <- function(curve_data, acronym, sex) {
  key_colours <- c("P03" = "#fa2c2e",
                   "P10" = "black",
                   "P25" = "#ff7d2b",
                   "P50" = "#09843b",
                   "P75" = "#ff7d2b",
                   "P90" = "black",
                   "P97" = "#fa2c2e")
  ggplot2::ggplot(
    curve_data, ggplot2::aes(x = gest_days, y = y, colour = Centile,
                             id = standard)) +
    ggplot2::geom_vline(xintercept = 168, colour = "grey60", linetype = "dashed") +
    ggplot2::geom_vline(xintercept = 300, colour = "grey60", linetype = "dashed") +
    ggplot2::geom_line() +
    ggplot2::scale_x_continuous(breaks = seq(154, 314, 14),
                                labels = seq(154, 314, 14)) +
    ggplot2::scale_colour_manual(values = key_colours) +
    ggplot2::labs(
      x = "Gestational age (days)",
      y = switch(acronym,
                 "wfga" = "Weight (kg)",
                 "lfga" = "Length (cm)",
                 "hcfga" = "Head circumference (cm)"))
}
```

#### Weight-for-GA: male

```{r ext_plot_wfga_male, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for male weight-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("wfga", "male") |>
  plot_extrapolated_chart("wfga", "male")
```

#### Weight-for-GA: female

```{r ext_plot_wfga_female, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for female weight-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("wfga", "female") |>
  plot_extrapolated_chart("wfga", "female")
```

#### Length-for-GA: male

```{r ext_plot_lfga_male, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for male length-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("lfga", "male") |>
  plot_extrapolated_chart("lfga", "male")
```

#### Length-for-GA: female

```{r ext_plot_lfga_female, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for female length-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("lfga", "female") |>
  plot_extrapolated_chart("lfga", "female")
```

#### Head circumference-for-GA: male

```{r ext_plot_hcfga_male, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for male head circumference-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("hcfga", "male") |>
  plot_extrapolated_chart("hcfga", "male")
```

#### Head circumference-for-GA: female

```{r ext_plot_hcfga_female, echo = FALSE, fig.alt = "A line graph showing the extrapolated growth curves for female head circumference-for-gestational age relative to the existing growth curves. Vertical light grey dashed lines indicate where extrapolation of the original INTERGROWTH-21st Newborn Size standards starts and stops. The extrapolated data (from 154 days GA to 168 days GA for the Very Preterm standards and from 301 to 314 days in the Newborn Size standards) appears to appropriately extend these standards past the existing GA bounds, at both the median (P50) and extreme centiles (P03 and P97)."}
generate_extrapolated_ig_nbs("hcfga", "female") |>
  plot_extrapolated_chart("hcfga", "female")
```

# Conclusion

In all, these extrapolated versions of the Newborn Size standards (incl. Very
Preterm Newborns) should be useful for researchers with large datasets that
contain relatively extreme gestational ages. Anyone utilising these extended
standards in their work should do so with care, and note specifically that
they used the extended versions.