---
title: "INTERGROWTH-21st Body Composition Equations"
description: >
  Deriving the INTERGROWTH-21<sup>st</sup> body composition equations.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{INTERGROWTH-21st Body Composition Equations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny", 
                      dev.args = list(png = list(type = "cairo")))
library(gigs)
fmfga_male <- tibble::as_tibble(gigs::ig_nbs$fmfga$male$centiles)
fmfga_female <- tibble::as_tibble(gigs::ig_nbs$fmfga$female$centiles)
bfpfga_male <- tibble::as_tibble(gigs::ig_nbs$bfpfga$male$centiles)
bfpfga_female <- tibble::as_tibble(gigs::ig_nbs$bfpfga$female$centiles)
ffmfga_male <- tibble::as_tibble(gigs::ig_nbs$ffmfga$male$centiles)
ffmfga_female <- tibble::as_tibble(gigs::ig_nbs$ffmfga$female$centiles)
```

Though Villar *et al.* (2017) published centiles for body composition (fat mass,
body fat percentage, and fat-free mass) in newborns from the
INTERGROWTH-21<sup>st</sup> project, they did not publish the parameters of the
models used to generate their centiles. To implement these standards in `gigs`,
we re-derived these values, in this case using the published data to work back
to linear equations.

## Contextual clues
To identify the way in which these body composition standards were modelled,
we need to look into how the published centile tables are formatted. Consider
the standard for fat mass for gestational age (`"fmfga"`) in females:
```{r looking_into_data_1}
fmfga_female
```

### Differences between each GA in `fmfga`
Let's first compare the differences over time in fat-mass for GA. In each
column, we'll subtract the value of the previous row from the current row to
produce a table of differences (see `check_differences()` below). Then we'll
check that the differences in each row are equal for all centiles, and
store this in the `all_equal` column.
```{r looking_into_data_2}
check_differences <- function(centile_tbl) {
  centile_tbl |>
    dplyr::mutate(dplyr::across(.cols = tidyselect::starts_with(match = "P"),
                                .fns = list(diff = \(x) x - dplyr::lag(x))),
                  .keep = "unused") |>
    dplyr::filter(dplyr::if_any(
            tidyselect::starts_with(match = "P"), ~ !is.na(.))) |>
    dplyr::rowwise() |>
    dplyr::mutate(
      all_equal = dplyr::n_distinct(P03_diff, P10_diff, P50_diff, P90_diff,
                                    P97_diff, na.rm = TRUE),
      all_equal = as.logical(all_equal)) |>
    dplyr::ungroup()
}
```

Let's check the results for `fmfga_female`:
```{r looking_into_data_3}
check_differences(fmfga_female)
```
The differences in fat mass are equal across percentiles at all gestational
ages, according to the `all_equal` column. If the standard deviation ($\sigma$)
changed with gestational age (GA), then `all_equal` would not be at each age.
This hints that for these centile curves, the mean value ($\mu$) for fat mass
was calculated first, then summed with the product of the desired
$z$-score/centile with a standard deviation ($\sigma$) which is constant
across the different GAs:
\begin{align*}
y = \mu + z\times\sigma
\end{align*}

#### In the other standards
Does this equality hold for the other centile tables?
```{r looking_into_data_4}
other_standards <- list(fmfga_male, bfpfga_male, bfpfga_female, ffmfga_male,
                        ffmfga_female)
vapply(X = other_standards,
       FUN = \(x) all(check_differences(x)$all_equal),
       FUN.VALUE = logical(length = 1L))
```
**No.** The first value, corresponding to `fmfga_male`, is `FALSE`. Let's see
what the differences look like for this table:
```{r looking_into_data_5}
check_differences(fmfga_male)
```
The issue here is a value of `38` rather than the `39`s in the `P10_diff`
column - but this difference is small enough to be a consequence of rounding, as
opposed to a real difference in $\sigma$ for this GA compared to the other GAs.

## Conclusion
**To reconstruct these standards, we need to determine the equations for $\mu$
in each standard. We also need to calculate $\sigma$, which does not change
with GA.**

# Extracting models from published tables
## Custom functions
The following custom functions perform several tasks:

* `run_LMs()` fits models to each INTERGROWTH-21<sup>st</sup> normative body
  composition standard using `lm()`, then returns the fitted models in a list.
* `std_deviation()` determines $\sigma$ for a given centile.
* `extract_SDs()` gets the standard deviations for each centile at each GA.
* `extract_eqn()` uses the above functions to extract a 'best possible' equation
  for the standard in use, then prints this to the console.
```{r custom_functions, echo = TRUE}
run_LMs <- function(df) {
  model_linear <- lm(P50 ~ gest_days, data = df)
  model_poly1 <- lm(P50 ~ poly(gest_days, degree = 1, raw = TRUE), data = df)
  model_poly2 <- lm(P50 ~ poly(gest_days, degree = 2, raw = TRUE), data = df)
  model_poly3 <- lm(P50 ~ poly(gest_days, degree = 3, raw = TRUE), data = df)
  list(model_linear, model_poly1, model_poly2, model_poly3)
}

#' Extract standard deviation from centiles
#' @param P50 Value of y at mean
#' @param PX Value of y at centile specified by `p`
#' @param p Centile at which y = PX
std_deviation <- function(P50, PX, p) (PX - P50) / qnorm(p = p)

extract_SDs <- function(df) {
  stddev_03 <- std_deviation(df$P50, df$P03, p = 0.03)
  stddev_10 <- std_deviation(df$P50, df$P10, p = 0.10)
  stddev_90 <- std_deviation(df$P50, df$P90, p = 0.90)
  stddev_97 <- std_deviation(df$P50, df$P97, p = 0.97)
  stddevs <- c(stddev_03, stddev_10, stddev_90, stddev_97)
  mean(stddevs, na.rm = TRUE)
}

#' Extract equation and standard deviation; print to console
#' @param df A centile table for `fmfga`, `bfpfga` or `ffmfga`.
extract_eqn <- function(df, verbose = TRUE) {
  li_models <- run_LMs(df)
  r_sq_adj <- vapply(X = li_models,
                     FUN = \(x) summary(x)$adj.r.squared,
                     FUN.VALUE = numeric(length = 1L))
  chosen_model <- li_models[[which(r_sq_adj == max(r_sq_adj))]]
  if (!verbose) {
    return(chosen_model)
  }
  model_coeffs <- coef(chosen_model)
  coef_str <- lapply(
    X = 2:length(model_coeffs),
    FUN = \(x) {
      start_str <- if (sign(model_coeffs[x]) == -1) "" else "+"
      end_str <- if (x == 2) "* x" else paste0("* x^", x - 1)
      paste(start_str, round(model_coeffs[x], 5), end_str)
    }) |>
    paste(collapse = " ")
  sign_ <- if (sign(model_coeffs[2]) == -1) "" else "+"
  cat(paste("y =", round(model_coeffs[[1]], 5), sign_, coef_str), sep = "\n")

  SDs <- extract_SDs(df)

  cat(paste("sigma =", round(mean(SDs, na.rm = TRUE), digits = 4)), sep = "\n")
  invisible(chosen_model)
}
```


## **Fat mass**
### Males
```{r fmfga_male_equation}
fmfga_male <- gigs::ig_nbs$fmfga$male$centiles
extract_eqn(fmfga_male)
```

### Females
```{r fmfga_female_equation}
fmfga_female <- gigs::ig_nbs$fmfga$female$centiles
extract_eqn(fmfga_female)
```

## **Body fat percentage**

### Males
```{r bfpfga_male_equation}
bfpfga_male <- gigs::ig_nbs$bfpfga$male$centiles
extract_eqn(bfpfga_male)
```

### Females

```{r bfpfga_female_equation}
bfpfga_female <- gigs::ig_nbs$bfpfga$female$centiles
extract_eqn(bfpfga_female)
```

## **Fat-free mass**
### Males

```{r ffmfga_male_equation}
ffmfga_male <- gigs::ig_nbs$ffmfga$male$centiles
extract_eqn(ffmfga_male)
```

### Females

```{r ffmfga_female_equation}
ffmfga_female <- gigs::ig_nbs$ffmfga$female$centiles
extract_eqn(ffmfga_female)
```

## Summary

### Equations for the mean:

| Standard      | Sex    | Equation for the mean ($\mu$)                                                                                                   |
|---------------|--------|---------------------------------------------------------------------------------------------------------------------------------|
| Fat mass      | Male   | $\mu = 96787.8 - 1056.77381\times\operatorname{GA} + 3.83673\times\operatorname{GA}^2 - 0.00462\times\operatorname{GA}^3$       |
| Fat mass      | Female | $\mu = 44432.51429 - 490.7602\times\operatorname{GA} + 1.80321\times\operatorname{GA}^2 - 0.00219\times\operatorname{GA}^3$     |
| Body fat %    | Male   | $\mu = 1521.84857 - 16.62398\times\operatorname{GA} + 0.0605\times\operatorname{GA}^2 - 0.00007\times\operatorname{GA}^3$       |
| Body fat %    | Female | $\mu = -66.09143 + 0.48102\times\operatorname{GA} - 0.00073\times\operatorname{GA}^2$                                           |
| Fat-free mass | Male   | $\mu = 347879.97143 - 3780.52721\times\operatorname{GA} + 13.73032\times\operatorname{GA}^2 - 0.01652\times\operatorname{GA}^3$ |
| Fat-free mass | Female | $\mu = 154457.42857 - 1689.0085\times\operatorname{GA} + 6.20845\times\operatorname{GA}^2 - 0.00753\times\operatorname{GA}^3$   |

### Standard deviations:
| Standard      | Sex    | Standard deviation ($\sigma$) |
|---------------|--------|-------------------------------|
| Fat mass      | Male   | $152.0728$                    |
| Fat mass      | Female | $156.8449$                    |
| Body fat %    | Male   | $3.6562$                      |
| Body fat %    | Female | $3.9547$                      |
| Fat-free mass | Male   | $276.5094$                    |
| Fat-free mass | Female | $261.0625$                    |

# Comparing fitted values with published values
All plotting code for this section can be found in the source code for this
vignette.

## Reconstructing the equations in R

First, we will bind the coefficients from each model into a table. This table
can then be used to look up the coefficients for each body composition standard.
```{r normative_bodycomp_functions, message = FALSE, warning = FALSE, error = FALSE}
all_standards <- list(fmfga_male, fmfga_female, bfpfga_male, bfpfga_female,
                      ffmfga_male, ffmfga_female)
models <- lapply(all_standards, FUN = \(standard) {
  model_coeffs <- standard |>
    extract_eqn(verbose = FALSE) |>
    coef() |>
    t() |>
    as.data.frame()
  if (ncol(model_coeffs) == 3) model_coeffs$extra <- 0
  names(model_coeffs) <- c("intercept", "x", "x^2", "x^3")
  model_coeffs$SD <- extract_SDs(standard)
  model_coeffs
}) |>
  do.call(what = "rbind")
rownames(models) <- c("fmfga_M" , "fmfga_F" , "bfpfga_M", "bfpfga_F",
                      "ffmfga_M", "ffmfga_F")
models
```

The `body_comp_centile2value()` will estimate `y` for a given gestational age
in days (`x`), `sex` and `acronym`. The `body_comp_centile2value_old()`
function does the same, but uses parameters from our older linear models for
each standard.
```{r body_comp_function}
ig_nbs_bc_centile2value <- function(p, x, sex, acronym) {
  params <- t(models)[, paste0(acronym, "_", sex)]
  SD <- params[[5]]
  mu <- params[[1]] + params[[2]] * x + params[[3]] * x^2 + params[[4]] * x^3
  mu + qnorm(p) * SD
}

ig_nbs_bc_centile2value_old <- function(p, x, sex, acronym) {
  z <- qnorm(p)
  if (sex == "M") {
    switch(acronym,
           "fmfga"  = -1134.2 + 37.2  * x/7 + z * 152.1593,
           "bfpfga" = -17.68  + 0.69  * x/7 + z * 3.6674,
           "ffmfga" = -2487.6 + 139.9 * x/7 + z * 276.2276,
           stop("Bad acronym", .call = F))
  } else if (sex == "F") {
    switch(acronym,
           "fmfga"  = -840.2 + 30.7  * x/7 + z * 156.8411,
           "bfpfga" = -9.02  + 0.51  * x/7 + z * 3.9405,
           "ffmfga" = -1279  + 105.3 * x/7 + z * 260.621,
           stop("Bad acronym", .call = F))
  }
}
```

# Comparing fitted values with published values
## Making percentile tables
```{r repl_fns}
#' Replicate centile tables as presented by Villar et al.
repl_centile_tbls <- function(sex, acronym, bodycomp_fn) {
  centiles <- c(0.03, 0.1, 0.5, 0.9, 0.97) |>
    purrr::set_names(c("P03", "P10", "P50", "P90", "P97"))
  tbl <- purrr::map(.x = centiles, .f = ~ {
      p <- .x
      column <- seq(266, 294, by = 7) |>
        purrr::set_names(paste(1:5)) |>
        purrr::map_dfc(.f  = ~ {
          bodycomp_fn(p, .x, sex, acronym)
        }) |>
      t()
    column[which(column <= 0)] <- NA_real_
    as.data.frame(column)
    })
  tbl <- suppressMessages(suppressWarnings(dplyr::bind_cols(tbl)))
  colnames(tbl) <- names(centiles)
  dplyr::mutate(tbl, gest_days = (38:42 * 7), .before = P03)
}

#' Make centile tables based on parameters from new polynomial models
repl_centile_tbls_new <- function(sex, acronym) {
  repl_centile_tbls(sex, acronym, ig_nbs_bc_centile2value)
}

#' Make centile tables based on parameters from linear models
repl_centile_tbls_old <- function(sex, acronym) {
  repl_centile_tbls(sex, acronym, ig_nbs_bc_centile2value_old)
}
```

```{r plotting_functions, echo = FALSE}
comparison_plot <- function(obs_tbl, pred_tbl, title, xlab, ylab, legend.pos = "none") {
  tbl <- dplyr::bind_rows(obs_tbl, pred_tbl) |>
    dplyr::mutate(predicted = c(rep("Observed", 5), rep("Fitted", 5))) |>
    tidyr::pivot_longer(cols = tidyselect::contains("P", ignore.case = F),
                        names_to = "Fitted", values_to = "yval") |>
    dplyr::filter(!is.na(yval))
  suppressWarnings(suppressMessages(
    plot <- ggplot2::ggplot(data = tbl, ggplot2::aes(x = gest_days, y = yval)) +
      ggplot2::geom_point(data = subset(tbl, predicted == "Observed"),
                          ggplot2::aes(shape = as.factor(1)),
                          colour = "red", fill = NA,
                          stroke = 1.3,
                          size = 1.5) +
      ggplot2::geom_line(data = subset(tbl, predicted == "Fitted"),
                         ggplot2::aes(colour = Fitted),
                         linetype = "dashed", size = 1.1) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      ggplot2::scale_shape_manual(name = "", values = c(1, rep(NA, 5)),
                                  labels = c("Observed", rep(NA, 5))) +
      ggplot2::scale_linetype(name = "") +#, values = c(NA, rep("dashed", 5))) +
      viridis::scale_color_viridis(name = "",
                                   labels = c("Fitted - P03", "Fitted - P10",
                                              "Fitted - P50", "Fitted - P90",
                                              "Fitted - P97"),
                                   discrete = T, option = "turbo") +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos) +
      ggplot2::guides(colour = ggplot2::guide_legend(order = 2),
                      shape = ggplot2::guide_legend(order = 1))
    ))
}

altman_esque_plot <- function(diff_tbl, title, xlab, ylab, legend.pos = "none") {
  diff_tbl_long <- diff_tbl |>
    dplyr::mutate(gest_days = (38:42 * 7)) |>
    tidyr::pivot_longer(cols = tidyselect::contains("P", ignore.case = F),
                        names_to = "Percentile", values_to = "yval") |>
    dplyr::mutate(gest_days = purrr::map2_vec(
      .x = gest_days,
      .y = Centile,
      .f = \(x,y) {
        x + switch(y, "P03" = 2, "P10" = 1, "P50" = 0, "P90" = 1, "P97" = 2)
      })) |>
    dplyr::filter(!is.na(yval))
  suppressWarnings(suppressMessages(
    plot <- ggplot2::ggplot(data = diff_tbl_long,
                            ggplot2::aes(x = gest_days, y = yval)) +
      ggplot2::geom_point(ggplot2::aes(colour = Centile),
                          shape = as.factor(x = 1),
                          fill = NA,
                          stroke = 1.3,
                          size = 1.5) +
      ggplot2::geom_hline(yintercept = 0, linetype = "dashed", size = 1.1) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      viridis::scale_color_viridis(name = "Centile",
                                   labels = c("Fitted - P03", "Fitted - P10",
                                              "Fitted - P50", "Fitted - P90",
                                              "Fitted - P97"),
                                   discrete = T, option = "turbo") +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos)
    ))
}

bland_altman_plot <- function(actual, predicted, title, xlab, ylab, legend.pos = "none") {
  actual_long <- actual |>
    tidyr::pivot_longer(cols = tidyselect::starts_with(match = "P")) |>
    dplyr::rename(reference = value)
  predicted_long <- predicted |>
    tidyr::pivot_longer(cols = tidyselect::starts_with(match = "P")) |>
    dplyr::rename(predicted = value)
  ba_tbl <- dplyr::full_join(actual_long, predicted_long) |>
    dplyr::rename(centile = name) |>
    dplyr::mutate(average = (reference + predicted) / 2,
                  difference = reference - predicted)
  hlinesize <- 1.1
  labelsize <- 2.5
  plot <- ggplot2::ggplot(data = ba_tbl,
                            ggplot2::aes(x = average, y = difference)) +
      ggplot2::geom_point(alpha = 0.5) +

      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T)),
                          linetype = "dashed", colour = "black", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T),
                                       label = paste("Mean = ",
                                                     round(mean(difference,
                                                                na.rm = T),
                                                           digits = 3))),
                         hjust = 0, colour = "black", fontface = "bold", size = labelsize) +
      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE)),
                          linetype = "dashed", colour = "red", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE),
                                       label = paste("Mean + 1.96 SD = ",
                                                     round(mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE),
                                                           digits = 3))),
                          hjust = 0, colour = "red", fontface = "bold", size = labelsize) +
      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE)),
                          linetype = "dashed", colour = "red", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE),
                                       label = paste("Mean - 1.96 SD = ",
                                                     round(mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE),
                                                           digits = 3))),
                          hjust = 0, colour = "red", fontface = "bold", size = labelsize) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos)
}
```

```{r fmfga_plots, echo = F, warning = FALSE, message = FALSE}
fmfga_pred_male <- repl_centile_tbls_new(sex = "M", acronym = "fmfga")
fmfga_male_p1 <- comparison_plot(obs_tbl = fmfga_male , pred_tbl = fmfga_pred_male,
                                 xlab = "Gestational age (days)", ylab = "Fat mass (g)",
                                 title = "", legend.pos = "bottom")
fmfga_male_p2 <- bland_altman_plot(actual = fmfga_male, predicted = fmfga_pred_male,
                                   xlab = "", ylab = "Difference in\nmeasurements (g)",
                                   title = "")
fmfga_male_p3 <- bland_altman_plot(actual = fmfga_male, predicted = repl_centile_tbls_old(sex = "M", acronym = "fmfga"),
                                   xlab = "", ylab = "Difference in\nmeasurements (g)",
                                   title = "")

fmfga_pred_female <- repl_centile_tbls_new(sex = "F", acronym = "fmfga")
fmfga_female_p1 <- comparison_plot(obs_tbl = fmfga_female , pred_tbl = fmfga_pred_female,
                      xlab = "Gestational age (days)", ylab = "",
                      title = "")
fmfga_female_p2 <- bland_altman_plot(actual = fmfga_female, predicted = fmfga_pred_female,
                                     xlab = "", ylab = "",
                                     title = "")
fmfga_female_p3 <- bland_altman_plot(actual = fmfga_male, predicted = repl_centile_tbls_old(sex = "F", acronym = "fmfga"),
                                     xlab = "Gestational age (days)", ylab = "",
                                     title = "")
```

```{r bfpfga_plots, echo = F, warning = FALSE, message = FALSE}
bfpfga_pred_male <- repl_centile_tbls_new(sex = "M", acronym = "bfpfga")
bfpfga_male_p1 <- comparison_plot(obs_tbl = bfpfga_male , pred_tbl = bfpfga_pred_male,
                      xlab = "", ylab = "% Body fat",
                      title = "")
bfpfga_male_p2 <- bland_altman_plot(actual = bfpfga_male, predicted = bfpfga_pred_male,
                                     xlab = "", ylab = "Difference in\nmeasurements (%)",
                                     title = "")
bfpfga_male_p3 <- bland_altman_plot(actual = bfpfga_male, predicted = repl_centile_tbls_old(sex = "M", acronym = "bfpfga"),
                                    xlab = "", ylab = "Difference in\nmeasurements (%)",
                                    title = "")

bfpfga_pred_female <- repl_centile_tbls_new(sex = "F", acronym = "bfpfga")
bfpfga_female_p1 <- comparison_plot(obs_tbl = bfpfga_female , pred_tbl = bfpfga_pred_female,
                      xlab = "", ylab = "% Body fat",
                      title = "")
bfpfga_female_p2 <- bland_altman_plot(actual = bfpfga_female, predicted = bfpfga_pred_female,
                                     xlab = "", ylab = "",
                                     title = "")
bfpfga_female_p3 <- bland_altman_plot(actual = bfpfga_male, predicted = repl_centile_tbls_old(sex = "F", acronym = "bfpfga"),
                                      xlab = "", ylab = "",
                                      title = "")
```

```{r ffmfga_plots, echo = F, warning = FALSE, message = FALSE}
ffmfga_pred_male <- repl_centile_tbls_new(sex = "M", acronym = "ffmfga")
ffmfga_male_p1 <- comparison_plot(obs_tbl = ffmfga_male , pred_tbl = ffmfga_pred_male,
                      xlab = "Gestational age (days)", ylab = "Fat-free mass (g)",
                      title =  "")
ffmfga_male_p2 <- bland_altman_plot(actual = ffmfga_male, predicted = ffmfga_pred_male,
                                    xlab = "Average measurement", ylab = "Difference in\nmeasurements (g)",
                                    title = "")
ffmfga_male_p3 <- bland_altman_plot(actual = ffmfga_male, predicted = repl_centile_tbls_old(sex = "M", acronym = "ffmfga"),
                                    xlab = "Average measurement", ylab = "Difference in\nmeasurements (g)",
                                    title = "")

ffmfga_pred_female <- repl_centile_tbls_new(sex = "F", acronym = "ffmfga")
ffmfga_female_p1 <- comparison_plot(obs_tbl = ffmfga_female , pred_tbl = ffmfga_pred_female,
                      xlab = "Gestational age (days)", ylab = "",
                      title = "")
ffmfga_female_p2 <- bland_altman_plot(actual = ffmfga_female, predicted = ffmfga_pred_female,
                                      xlab = "Average measurement", ylab = "",
                                      title = "")
ffmfga_female_p3 <- bland_altman_plot(actual = ffmfga_female, predicted = repl_centile_tbls_old(sex = "F", acronym = "ffmfga"),
                                      xlab = "Average measurement", ylab = "",
                                      title = "")
```

## Observed vs fitted values

```{r comparison_plots, echo = F, fig.width = 8, fig.height = 12, fig.cap = "Observed and estimated normative centiles for body composition measures according to gestational age and sex. Centiles (3rd, 10th, 50th, 90th, and 97th) for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to gestational age. Red circles show the observed value for that centile and gestational age; dashed lines show estimates from linear equations fitted to these observed values.", warning = FALSE, message = FALSE, fig.alt = "Observed and estimated normative centiles for body composition measures according to gestational age and sex. Centiles (3rd, 10th, 50th, 90th, and 97th) for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to gestational age. Red circles show the observed value for that centile and gestational age; dashed lines show estimates from linear equations fitted to these observed values."}
overall_legend <- cowplot::get_legend(fmfga_male_p1)
fmfga_male_p1_noleg <- fmfga_male_p1 + ggplot2::theme(legend.position = "none")
plotlist <- list(fmfga_male_p1_noleg, fmfga_female_p1, bfpfga_male_p1, bfpfga_female_p1, ffmfga_male_p1, ffmfga_female_p1)
cp <- cowplot::plot_grid(plotlist = plotlist, labels = "auto", ncol = 2, rel_heights = 5, rel_widths = 0.5)
cowplot::plot_grid(plotlist = list(cp, overall_legend), labels = NULL, ncol = 1, rel_heights = c(1, .1))
```

## Bland-Altman plots
### Reference vs. predicted values
```{r bland_altman_plots, echo = F, fig.width = 8, fig.height = 12, fig.cap = "Bland-Altman plots for observed and estimated centile values for body composition measures according to gestational age and sex. Circles show differences between observed and estimated values for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to the average of the two measurements.", warning = FALSE, message = FALSE, fig.alt = "Observed and estimated normative centiles for body composition measures according to gestational age and sex. Centiles (3rd, 10th, 50th, 90th, and 97th) for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to gestational age. Red circles show the observed value for that centile and gestational age; dashed lines show estimates from linear equations fitted to these observed values."}
plotlist <- list(fmfga_male_p2, fmfga_female_p2, bfpfga_male_p2, bfpfga_female_p2, ffmfga_male_p2, ffmfga_female_p2)
cowplot::plot_grid(plotlist = plotlist, labels = "auto", ncol = 2, rel_heights = 5, rel_widths = 0.5)
```


# References
Villar J, Puglia FA, Fenton TR, Ismal LC, Staines-Urias E, Giuliani F, *et al.*
**Body composition at birth and its relationship with neonatal anthropometric
ratios: the newborn body composition study of the INTERGROWTH-21st project**
*Pediatric Research* 2017, **82:305-316.** doi:
[10.1038/pr.2017.52](https://doi.org/10.1038/pr.2017.52)
