---
title: "INTERGROWTH-21st Body Composition Equations: Improved"
description: >
  Our improved process for deriving the INTERGROWTH-21<sup>st</sup> body composition equations.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{INTERGROWTH-21st Body Composition Equations: Improved}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny", dev.args = list(png = list(type = "cairo")))
library(gigs)
```

Though Villar *et al.* (2017) published centiles for body composition (fat mass, body fat percentage, and fat-free
mass) in newborns from the INTERGROWTH-21<sup>st</sup> project, they did not publish the parameters of the models used
to generate their centiles. To implement these standards in our package we re-derives these values, in this case using
the published data to work back to linear equations for the median and SD of each standard.

# Extracting models from published data

## Custom functions
The following custom functions fit linear models to each INTERGROWTH-21<sup>st</sup> body composition standard, then
print the results:
```{r custom_functions, echo = TRUE}
#' Extract standard deviation from percentiles
#' @param P50 Value of y at mean
#' @param PX Value of y at percentile specified by p
#' @param p Percentile at which y = PX
std_deviation <- function(P50, PX, p) (PX - P50) / qnorm(p = p)

#' Model equation for median (50th centile); print to console
#' @param df A dataframe with columns gest_age and P50. Used to fit LM and
#' get standard deviation.
LM_median <- function(df) {
  model <- lm(P50 ~ gest_age, data = df)
  std_dev <- mapply(FUN = std_deviation,
                     list(df$P50), list(df$P03, df$P10, df$P90, df$P97),
                     list(0.03, 0.10, 0.90, 0.97)) |>
    median(na.rm = TRUE)
  str_model <- paste("y =", coef(model)[[1]], "+",
                     coef(model)[[2]], "* x", "+",
                     "z * ", std_dev)
  str_model
}
```

```{r, commented_out, include = FALSE}
# #' Extract equation and standard deviation; print to console
# #' @param df A dataframe with columns gest_age, P50, and P90. Used to fit LM and
# #' get standard deviation.
# LM_stddev <- function(df) {
#   std_devs <- mapply(FUN = std_deviation,
#                      list(df$P50), list(df$P03, df$P10, df$P90, df$P97),
#                      list(0.03, 0.10, 0.90, 0.97))
#   mean(std_devs, na.rm = T)
#   names(std_devs) <- c("sigma_03","sigma_10","sigma_90","sigma_97")
#   df <- cbind(df, std_devs)
#   # out <- list(df, model_03, model_10, model_90, model_97)
#   # names(out) <- c("df", "sigma_03","sigma_10","sigma_90","sigma_97")
#   # invisible(out)
#   mean(std_devs, na.rm = T)
# }
#
#
# #' Extract equation and standard deviation; print to console
# #' @param df A dataframe with columns gest_age, P50, and P90. Used to fit LM and
# #' get standard deviation.
# LM_stddev2 <- function(df) {
#   std_devs <- lapply(FUN = function(x) {
#     temp <- df[x, 2:6]
#     temp |>
#       t() |>
#       sd(na.rm = T)
#   }, seq_along(df)) #list(df[1, 1:5], df[2, 1:5]))
#   std_devs
# }

```

## **Fat mass**
### Males
```{r fmfga_male_equation}
fmfga_male <- gigs::ig_nbs$fmfga$male$percentiles
LM_median(fmfga_male)
# LM_stddev(fmfga_male)
# t <- LM_stddev2(fmfga_male)
```

### Females
```{r fmfga_female_equation}
fmfga_female <- gigs::ig_nbs$fmfga$female$percentiles
LM_median(fmfga_female)
```

## **Body fat percentage**

### Males
```{r bfpfga_male_equation}
bfpfga_male <- gigs::ig_nbs$bfpfga$male$percentiles
LM_median(bfpfga_male)
```

### Females

```{r bfpfga_female_equation}
bfpfga_female <- gigs::ig_nbs$bfpfga$female$percentiles
LM_median(bfpfga_female)
```

## **Fat-free mass**
### Males

```{r ffmfga_male_equation}
ffmfga_male <- gigs::ig_nbs$ffmfga$male$percentiles
LM_median(ffmfga_male)
```

### Females

```{r ffmfga_female_equation}
ffmfga_female <- gigs::ig_nbs$ffmfga$female$percentiles
LM_median(ffmfga_female)
```

## Summary

### Fat mass:

| Standard      | Sex    | Equation                                                                                  |
|---------------|--------|-------------------------------------------------------------------------------------------|
| Fat mass      | Male   | $\operatorname{fat\_mass} = -1134.2 + 37.2 (\operatorname{GA}) + z \times 152.063$        |
| Fat mass      | Female | $\operatorname{fat\_mass} = -840.2 + 30.7 (\operatorname{GA}) + z \times 156.844$         |
| Body fat %    | Male   | $\operatorname{body\_fat\_\%} = -17.68 + 0.69 (\operatorname{GA}) + z \times 3.6674$      |
| Body fat %    | Female | $\operatorname{body\_fat\_\%} = -9.02 + 0.51 (\operatorname{GA}) + z \times 3.9795$       |
| Fat-free mass | Male   | $\operatorname{fat\_free\_mass} = -2487.6 + 139.9 (\operatorname{GA}) + z \times 276.479$ |
| Fat-free mass | Female | $\operatorname{fat\_free\_mass} = -1279 + 105.3 (\operatorname{GA}) + z \times 261.060$   |

# Comparing fitted values with published values
All plotting code for this section can be found in the source code for this vignette.
```{r repl_fns, echo = FALSE}
#' Replicate centile tables as presented by Villar et al.
repl_centile_tbls <- function (sex, acronym) {
  centiles <- c(0.03, 0.1, 0.5, 0.9, 0.97) |>
    purrr::set_names(c("P03", "P10", "P50", "P90", "P97"))
  tbl <- purrr::map(.x = centiles, .f = ~ {
      p <- .x
      col <- 38:42 |>
        purrr::set_names(paste(1:5)) |>
        purrr::map_dfc(.f  = ~ {
          GA <- .x
          z <- qnorm(p)
          if (sex == "M") {
            switch(acronym,
                   "fmfga"  = -1134.2 + 37.2  * GA + z * 152.063,
                   "bfpfga" = -17.68  + 0.69  * GA + z * 3.6674,
                   "ffmfga" = -2487.6 + 139.9 * GA + z * 276.479,
                   stop("Bad acronym", .call = F))
          } else if (sex == "F") {
            switch(acronym,
                   "fmfga"  = -840.2 + 30.7  * GA + z * 156.844,
                   "bfpfga" = -9.02  + 0.51  * GA + z * 3.9795,
                   "ffmfga" = -1279  + 105.3 * GA + z * 261.060,
                   stop("Bad acronym", .call = F))
          }
        }) |> t()
    col[which(col <= 0)] <- NA_real_
    col <- as.data.frame(col)
    return(col)
    })
  tbl <- suppressMessages(suppressWarnings(dplyr::bind_cols(tbl)))
  colnames(tbl) <- names(centiles)
  dplyr::mutate(tbl, gest_age = (38:42 * 7), .before = P03)
}

comparison_plot <- function(obs_tbl, pred_tbl, title, xlab, ylab, legend.pos = "none") {
  tbl <- dplyr::bind_rows(obs_tbl, pred_tbl) |>
    dplyr::mutate(predicted = c(rep("Observed", 5), rep("Fitted", 5))) |>
    tidyr::pivot_longer(cols = tidyselect::contains("P", ignore.case = F),
                        names_to = "Fitted", values_to = "yval") |>
    dplyr::filter(!is.na(yval))
  suppressWarnings(suppressMessages(
    plot <- ggplot2::ggplot(data = tbl, ggplot2::aes(x = gest_age, y = yval)) +
      ggplot2::geom_point(data = subset(tbl, predicted == "Observed"),
                          ggplot2::aes(shape = as.factor(1)),
                          colour = "red", fill = NA,
                          stroke = 1.3,
                          size = 1.5) +
      ggplot2::geom_line(data = subset(tbl, predicted == "Fitted"),
                         ggplot2::aes(colour = Fitted),
                         linetype = "dashed", size = 1.1) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      ggplot2::scale_shape_manual(name = "", values = c(1, rep(NA, 5)),
                                  labels = c("Observed", rep(NA, 5))) +
      ggplot2::scale_linetype(name = "") +#, values = c(NA, rep("dashed", 5))) +
      viridis::scale_color_viridis(name = "",
                                   labels = c("Fitted - P03", "Fitted - P10",
                                              "Fitted - P50", "Fitted - P90",
                                              "Fitted - P97"),
                                   discrete = T, option = "turbo") +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos) +
      ggplot2::guides(colour = ggplot2::guide_legend(order = 2),
                      shape = ggplot2::guide_legend(order = 1))
    ))
}

altman_esque_plot <- function(diff_tbl, title, xlab, ylab, legend.pos = "none") {
  diff_tbl_long <- diff_tbl |>
    dplyr::mutate(gest_age = (38:42 * 7)) |>
    tidyr::pivot_longer(cols = tidyselect::contains("P", ignore.case = F),
                        names_to = "Percentile", values_to = "yval") |>
    dplyr::mutate(gest_age = purrr::map2_vec(.x = gest_age,
                                         .y = Percentile,
                                         .f = ~ .x + switch(.y,
                                               "P03" = -2,
                                               "P10" = -1,
                                               "P50" = 0,
                                               "P90" = 1,
                                               "P97" = 2))) |>
    dplyr::filter(!is.na(yval))
  suppressWarnings(suppressMessages(
    plot <- ggplot2::ggplot(data = diff_tbl_long,
                            ggplot2::aes(x = gest_age, y = yval)) +
      ggplot2::geom_point(ggplot2::aes(colour = Percentile),
                          shape = as.factor(1),
                          fill = NA,
                          stroke = 1.3,
                          size = 1.5) +
      ggplot2::geom_hline(yintercept = 0, linetype = "dashed", size = 1.1) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      viridis::scale_color_viridis(name = "Percentile",
                                   labels = c("Fitted - P03", "Fitted - P10",
                                              "Fitted - P50", "Fitted - P90",
                                              "Fitted - P97"),
                                   discrete = T, option = "turbo") +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos)
    ))
}

bland_altman_plot <- function(actual, predicted, title, xlab, ylab, legend.pos = "none") {
  actual_long <- actual |>
    tidyr::pivot_longer(cols = tidyselect::starts_with(match = "P")) |>
    dplyr::rename(reference = value)
  predicted_long <- predicted |>
    tidyr::pivot_longer(cols = tidyselect::starts_with(match = "P")) |>
    dplyr::rename(predicted = value)
  ba_tbl <- dplyr::full_join(actual_long, predicted_long) |>
    dplyr::rename(centile = name) |>
    dplyr::mutate(average = (reference + predicted) / 2,
                  difference = reference - predicted)
  hlinesize <- 1.1
  labelsize <- 2.5
  plot <- ggplot2::ggplot(data = ba_tbl,
                            ggplot2::aes(x = average, y = difference)) +
      ggplot2::geom_point(alpha = 0.5) +

      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T)),
                          linetype = "dashed", colour = "black", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T),
                                       label = paste("Mean = ",
                                                     round(mean(difference,
                                                                na.rm = T),
                                                           digits = 3))),
                         hjust = 0, colour = "black", fontface = "bold", size = labelsize) +
      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE)),
                          linetype = "dashed", colour = "red", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE),
                                       label = paste("Mean + 1.96 SD = ",
                                                     round(mean(difference, na.rm = T) + 1.96 * sd(difference, na.rm = TRUE),
                                                           digits = 3))),
                          hjust = 0, colour = "red", fontface = "bold", size = labelsize) +
      ggplot2::geom_hline(ggplot2::aes(yintercept = mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE)),
                          linetype = "dashed", colour = "red", size = hlinesize) +
      ggplot2::geom_label(ggplot2::aes(x = -Inf,
                                       y = mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE),
                                       label = paste("Mean - 1.96 SD = ",
                                                     round(mean(difference, na.rm = T) - 1.96 * sd(difference, na.rm = TRUE),
                                                           digits = 3))),
                          hjust = 0, colour = "red", fontface = "bold", size = labelsize) +
      ggplot2::labs(title = ggplot2::element_text(title),
                    x = ggplot2::element_text(xlab),
                    y = ggplot2::element_text(ylab)) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = legend.pos)
}
```

```{r fmfga_plots, echo = F, warning = FALSE, message = FALSE}
fmfga_pred_male <- repl_centile_tbls(sex = "M", acronym = "fmfga")
fmfga_male_p1 <- comparison_plot(obs_tbl = fmfga_male , pred_tbl = fmfga_pred_male,
                                 xlab = "Gestational age (days)", ylab = "Fat mass (g)",
                                 title = "", legend.pos = "bottom")
fmfga_male_p2 <- altman_esque_plot(diff_tbl = fmfga_pred_male - fmfga_male,
                                   xlab = "", ylab = "Difference from\nobserved (g)",
                                   title = "", legend.pos = "bottom")
fmfga_male_p3 <- bland_altman_plot(actual = fmfga_male, predicted = fmfga_pred_male,
                                   xlab = "", ylab = "Difference in\nmeasurements (%)",
                                   title = "")

fmfga_pred_female <- repl_centile_tbls(sex = "F", acronym = "fmfga")
fmfga_female_p1 <- comparison_plot(obs_tbl = fmfga_female , pred_tbl = fmfga_pred_female,
                      xlab = "Gestational age (days)", ylab = "",
                      title = "")
fmfga_female_p2 <- altman_esque_plot(diff_tbl = fmfga_pred_female - fmfga_female,
                        xlab = "", ylab = "",
                        title = "")
fmfga_female_p3 <- bland_altman_plot(actual = fmfga_female, predicted = fmfga_pred_female,
                                     xlab = "", ylab = "",
                                     title = "")
```

```{r bfpfga_plots, echo = F, warning = FALSE, message = FALSE}
bfpfga_pred_male <- repl_centile_tbls(sex = "M", acronym = "bfpfga")
bfpfga_male_p1 <- comparison_plot(obs_tbl = bfpfga_male , pred_tbl = bfpfga_pred_male,
                      xlab = "", ylab = "% Body fat",
                      title = "")
bfpfga_male_p2 <- altman_esque_plot(diff_tbl = bfpfga_pred_male - bfpfga_male,
                        xlab = "", ylab = "Difference from\nobserved (%)",
                        title = "")
bfpfga_male_p3 <- bland_altman_plot(actual = bfpfga_male, predicted = bfpfga_pred_male,
                                     xlab = "", ylab = "Difference in\nmeasurements (%)",
                                     title = "")

bfpfga_pred_female <- repl_centile_tbls(sex = "F", acronym = "bfpfga")
bfpfga_female_p1 <- comparison_plot(obs_tbl = bfpfga_female , pred_tbl = bfpfga_pred_female,
                      xlab = "", ylab = "% Body fat",
                      title = "")
bfpfga_female_p2 <- altman_esque_plot(diff_tbl = bfpfga_pred_female - bfpfga_female,
                        xlab = "", ylab = "",
                        title = "")
bfpfga_female_p3 <- bland_altman_plot(actual = bfpfga_female, predicted = bfpfga_pred_female,
                                     xlab = "", ylab = "",
                                     title = "")
```

```{r ffmfga_plots, echo = F, warning = FALSE, message = FALSE}
ffmfga_pred_male <- repl_centile_tbls(sex = "M", acronym = "ffmfga")
ffmfga_male_p1 <- comparison_plot(obs_tbl = ffmfga_male , pred_tbl = ffmfga_pred_male,
                      xlab = "Gestational age (days)", ylab = "Fat-free mass (g)",
                      title =  "")
ffmfga_male_p2 <- altman_esque_plot(diff_tbl = ffmfga_pred_male - ffmfga_male,
                        xlab = "Gestational age (days)", ylab = "Difference from\nobserved (g)",
                        title = "")
ffmfga_male_p3 <- bland_altman_plot(actual = ffmfga_male, predicted = ffmfga_pred_male,
                                    xlab = "Average measurement (g)", ylab = "Difference in\nmeasurements (g)",
                                    title = "")

ffmfga_pred_female <- repl_centile_tbls(sex = "F", acronym = "ffmfga")
ffmfga_female_p1 <- comparison_plot(obs_tbl = ffmfga_female , pred_tbl = ffmfga_pred_female,
                      xlab = "Gestational age (days)", ylab = "",
                      title = "")
ffmfga_female_p2 <- altman_esque_plot(diff_tbl = ffmfga_pred_female - ffmfga_female,
                        xlab = "Gestational age (days)", ylab = "",
                        title = "")
ffmfga_female_p3 <- bland_altman_plot(actual = ffmfga_female, predicted = ffmfga_pred_female,
                                      xlab = "Average measurement (g)", ylab = "",
                                      title = "")
```

## Observed vs fitted values

```{r comparison_plots, echo = F, fig.width = 8, fig.height = 12, fig.cap = "Observed and estimated normative centiles for body composition measures according to gestational age and sex. Centiles (3rd, 10th, 50th, 90th, and 97th) for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to gestational age. Red circles show the observed value for that centile and gestational age; dashed lines show estimates from linear equations fitted to these observed values."}
suppressWarnings( expr = {
  overall_legend <- cowplot::get_legend(fmfga_male_p1)
  fmfga_male_p1_noleg <- fmfga_male_p1 + ggplot2::theme(legend.position = "none")
  plotlist <- list(fmfga_male_p1_noleg, fmfga_female_p1, bfpfga_male_p1, bfpfga_female_p1, ffmfga_male_p1, ffmfga_female_p1)
  cp <- cowplot::plot_grid(plotlist = plotlist, labels = "auto", ncol = 2, rel_heights = 5, rel_widths = 0.5)
  cowplot::plot_grid(plotlist = list(cp, overall_legend), labels = NULL, ncol = 1, rel_heights = c(1, .1))
})
```

## Difference between observed and fitted values
```{r altman_esque_plots, echo = F, fig.width = 8, fig.height = 12, fig.cap = "Difference between observed and estimated centile values for body composition measures according to gestational age and sex. Circles show differences between observed and estimated values for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to gestational age, coloured by centiles (3rd, 10th, 50th, 90th and 97th)."}
suppressWarnings( expr = {
  overall_legend <- cowplot::get_legend(fmfga_male_p2)
  fmfga_male_p2_noleg <- fmfga_male_p2 + ggplot2::theme(legend.position = "none")
  plotlist <- list(fmfga_male_p2_noleg, fmfga_female_p2, bfpfga_male_p2, bfpfga_female_p2, ffmfga_male_p2, ffmfga_female_p2)
  cp <- cowplot::plot_grid(plotlist = plotlist, labels = "auto", ncol = 2, rel_heights = 5, rel_widths = 0.5)
  cowplot::plot_grid(plotlist = list(cp, overall_legend), labels = NULL, ncol = 1, rel_heights = c(1, .1))
})
```

## Bland-Altman plots for reference and predicted values
```{r bland_altman_plots, echo = F, fig.width = 8, fig.height = 12, fig.cap = "Bland-Altman plots for observed and estimated centile values for body composition measures according to gestational age and sex. Circles show differences between observed and estimated values for fat mass (**a** males; **b** females); body fat percentage (**c** males; **d** females); and fat-free mass (**e** males; **f** females) according to the average of the two measurements."}
suppressWarnings( expr = {
  plotlist <- list(fmfga_male_p3, fmfga_female_p3, bfpfga_male_p3, bfpfga_female_p3, ffmfga_male_p3, ffmfga_female_p3)
  cowplot::plot_grid(plotlist = plotlist, labels = "auto", ncol = 2, rel_heights = 5, rel_widths = 0.5)
})
```

# References

Villar J, Puglia FA, Fenton TR, Ismal LC, Staines-Urias E, Giuliani F, *et al.* **Body composition at birth and
its relationship with neonatal anthropometric ratios: the newborn body composition study of the INTERGROWTH-21st
project** *Pediatric Research* 2017, **82:305-316.** doi: [10.1038/pr.2017.52](https://doi.org/10.1038/pr.2017.52)
